---
title: "Statistical Analysis of Manually Scored Aggressive and Affiliative Behavior"
author: "Hannah Anderson"
date: "2023-04-25"
output: pdf_document
editor_options:
  chunk_output_type: console
---

## Introduction

This file was created to analyze data collected during a series of experiments
conducted from March 30 to August 14, 2022 on the effects of fluctuating
visibility on the behavior of zebrafish (Danio rerio). All data analyzed in this
file was scored using Behavioral Observation Research Interactive Software
(BORIS; Friard & Gamba, 2016) by trained observers. For more information on the
experiments see the file the readme file located in this repository.

The specific behaviors analyzed here are chasing, lateral display and
aggressive contact (aggressive behaviors) and following (affiliative behavior).
For chasing, lateral display and following we scored both the duration of each
behavior and the number of occurrences. Aggressive contact was an instantaneous
behavior and we therefore only scored the number of occurrences.

~will add definitions for each behavior later~

Most of the analyses were conducted using linear mixed modesl with tank ID as a 
random factor to account for multiple recordings of each tank. For the analysis
divided by high/low visibility fluctuations in the acute clay and both dye
experiments we instead used linear models because there were insufficient levels
of tank ID to warrant a random effect.


## Setup

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

setwd("C://Users/hande/Documents/R/fluctuating-turbidity")
library(tidyverse)
library(glmmTMB)
library(emmeans)
library(car)
library(DHARMa)
library(performance)

#importing the base data
chron_clay <- read_csv("chronic_clay_BORIS_all.csv")
clay_dye_all <- read_csv("comparison_clay_dye_BORIS_all_data.csv")
dye_dye_all <- read_csv("comparison_clay_dye_BORIS_all_data.csv")
#also importing the aggregated behavioral data
chron_agg <- read_csv("chronic_clay_BORIS_agg.csv")
clay_dye_agg <- read_csv("comparison_clay_dye_BORIS_agg_data.csv")
dye_dye_agg <- read_csv("comparison_clay_dye_BORIS_agg_data.csv")

#re-classifying the recording variable as a character for use in the filter()
#function
clay_dye_all$recording <- as.character(clay_dye_all$recording)
clay_dye_agg$recording <- as.character(clay_dye_agg$recording)
dye_dye_all$recording <- as.character(dye_dye_all$recording)
dye_dye_agg$recording <- as.character(dye_dye_agg$recording)

#a list of the scored behaviors to be used in filtering the data by behavior
behavs <- c("chasing", "lat_display", "agg_contact", "following")


#sub-setting the data by recording for the direct comparisons and filtering by
#behavior for the individual behavior models
direct_claydye <- list()
direct_dyedye <- list()
for (i in 1:4) {
  direct_claydye[[i]] <- clay_dye_all %>%
                         filter(recording == "4") %>%
                         filter(behavior == behavs[[i]])
}
for (i in 1:4) {
  direct_dyedye[[i]] <- clay_dye_all %>%
                         filter(recording == "4") %>%
                         filter(behavior == behavs[[i]])
}
#and also sub-setting by recording for the aggregated behavior models
direct_claydye[[5]] <- clay_dye_agg %>%
                       filter(recording == "4") %>%
                       filter(behavior == "all_aggressive")
direct_claydye[[6]] <- clay_dye_agg %>%
                       filter(recording == "4") %>%
                       filter(behavior == "all")
direct_dyedye[[5]] <- clay_dye_agg %>%
                       filter(recording == "4") %>%
                       filter(behavior == "all_aggressive")
direct_dyedye[[6]] <- clay_dye_agg %>%
                       filter(recording == "4") %>%
                       filter(behavior == "all")
```


## Analyses for Experiment 1: Chronic (Two Week) Clay Exposure
#  Analyses divided by behavior and including all recordings

```{r chronic clay all}
#filtering the data by behavior in order to model each behavior separately and
#putting all of them in a list
chronic_list <- list()
for (i in 1:(length(behavs))){
  chronic_list[[i]] <- chron_clay %>%
                     filter(behavior == behavs[i])
}

#models of behavior counts across all recordings
chronic_chase_c <- glmmTMB(behav_count ~ treatment*sex + (1|tankID),
                       data = chronic_list[[1]], family = gaussian())
chronic_lat_c <- update(chronic_chase_c, . ~ ., data = chronic_list[[2]])
chronic_cont_c <- update(chronic_chase_c, . ~ ., data = chronic_list[[3]])
chronic_foll_c <- update(chronic_chase_c, . ~ ., data = chronic_list[[4]])

#checking the model assumptions
plot(simulateResiduals(fittedModel = chronic_chase_c)) #assumptions not met
plot(simulateResiduals(fittedModel = chronic_lat_c)) #assumptions not met
plot(simulateResiduals(fittedModel = chronic_cont_c)) #assumptions not met
plot(simulateResiduals(fittedModel = chronic_foll_c)) #assumptions not met

#transforming the data to meet model assumptions
chron_chase_log <- as.data.frame(chronic_list[[1]])
chron_chase_log$behav_count <- log(chron_chase_log$behav_count+1)
chron_lat_log <- as.data.frame(chronic_list[[2]])
chron_lat_log$behav_count <- log(chron_lat_log$behav_count+1)
chron_cont_log <- as.data.frame(chronic_list[[3]])
chron_cont_log$behav_count <- log(chron_cont_log$behav_count+1)
chron_foll_log <- as.data.frame(chronic_list[[4]])
chron_foll_log$behav_count <- log(chron_foll_log$behav_count+1)

#the tricky transformation (still not working)
test <- update(chronic_lat_c, . ~ ., family = poisson())
plot(simulateResiduals(fittedModel = test))

#the transformed models
chrolog_chase_c <- update(chronic_chase_c, . ~ ., data = chron_chase_log)
chrolog_lat_c <- update(chronic_lat_c, . ~ ., data = chron_lat_log)
chrolog_cont_c <- update(chronic_cont_c, . ~ ., data = chron_cont_log)
chrolog_foll_c <- update(chronic_foll_c, . ~ ., data = chron_foll_log)

#checking the transformed model assumptions
plot(simulateResiduals(fittedModel = chrolog_chase_c))
plot(simulateResiduals(fittedModel = chrolog_lat_c)) #better than untransformed, assumptions still not met
plot(simulateResiduals(fittedModel = chrolog_cont_c))
plot(simulateResiduals(fittedModel = chrolog_foll_c))

#analyzing the results
Anova(chrolog_chase_c)
#Anova(chrolog_lat_c)
Anova(chrolog_cont_c)
Anova(chrolog_foll_c)


#models of behavior durations across all relevant recordings
chronic_chase_d <- update(chronic_chase_c, duration_s ~ .)
chronic_lat_d <- update(chronic_lat_c, duration_s ~ .)
chronic_foll_d <- update(chronic_foll_c, duration_s ~ .)

#checking the model assumptions
plot(simulateResiduals(fittedModel = chronic_chase_d)) #assumptions not met
plot(simulateResiduals(fittedModel = chronic_lat_d)) #assumptions not met
plot(simulateResiduals(fittedModel = chronic_foll_d)) #assumptions not met

#transforming the data to meet model assumptions
chron_chase_log$duration_s <- log(chron_chase_log$duration_s)
chron_lat_log$duration_s <- log(chron_lat_log$duration_s)
chron_foll_log$duration_s <- log(chron_foll_log$duration_s)

#the transformed models
chrolog_chase_d <- update(chronic_chase_d, . ~ ., data = chron_chase_log)
chrolog_lat_d <- update(chronic_lat_d, . ~ ., data = chron_lat_log)
chrolog_foll_d <- update(chronic_foll_d, . ~ ., data = chron_foll_log)

#checking the transformed model assumptions
plot(simulateResiduals(fittedModel = chrolog_chase_d))
plot(simulateResiduals(fittedModel = chrolog_lat_d))
plot(simulateResiduals(fittedModel = chrolog_foll_d))

#analyzing the results
Anova(chrolog_chase_d)
Anova(chrolog_lat_d)
Anova(chrolog_foll_d)
```


## Analyses for Experiment 1: Chronic (Two Week) Clay Exposure
#  Analyses divided by behavior and by high and low visibility fluctuations

```{r chronic clay highlow}
#modelling only the recordings taken when the flux condition had high visibility

#filtering the data by behavior in order to model each behavior separately and
#putting all of them in a list
chronhi_list <- list()
for (i in 1:(length(behavs))){
  chronhi_list[[i]] <- high_chron %>%
                     filter(behavior == behavs[i])
}

#models of behavior counts across high visibility fluctuations
chronhi_chase_c <- glmmTMB(behav_count ~ treatment*sex + (1|tankID),
                       data = chronhi_list[[1]], family = gaussian())
chronhi_lat_c <- update(chronhi_chase_c, . ~ ., data = chronhi_list[[2]])
chronhi_cont_c <- update(chronhi_chase_c, . ~ ., data = chronhi_list[[3]])
chronhi_foll_c <- update(chronhi_chase_c, . ~ ., data = chronhi_list[[4]])

#checking the model assumptions
plot(simulateResiduals(fittedModel = chronhi_chase_c)) #assumptions not met
plot(simulateResiduals(fittedModel = chronhi_lat_c)) #assumptions not met
plot(simulateResiduals(fittedModel = chronhi_cont_c)) #assumptions not met
plot(simulateResiduals(fittedModel = chronhi_foll_c))

#transforming the data to meet model assumptions
chronhi_chase_log <- as.data.frame(chronhi_list[[1]])
chronhi_chase_log$behav_count <- log(chronhi_chase_log$behav_count+1)
chronhi_lat_log <- as.data.frame(chronhi_list[[2]])
chronhi_lat_log$behav_count <- log(chronhi_lat_log$behav_count+1)
chronhi_cont_log <- as.data.frame(chronhi_list[[3]])
chronhi_cont_log$behav_count <- log(chronhi_cont_log$behav_count+1)

#the transformed models
chrologhi_chase_c <- update(chronhi_chase_c, . ~ ., data = chronhi_chase_log)
chrologhi_lat_c <- update(chronhi_lat_c, . ~ ., data = chronhi_lat_log)
chrologhi_cont_c <- update(chronhi_cont_c, . ~ ., data = chronhi_cont_log)

#checking the assumptions of the transformed models
plot(simulateResiduals(fittedModel = chrologhi_chase_c))
plot(simulateResiduals(fittedModel = chrologhi_lat_c))
plot(simulateResiduals(fittedModel = chrologhi_cont_c))

#analyzing the results
Anova(chrologhi_chase_c)
#Anova(chrologhi_lat_c)
Anova(chrologhi_cont_c)
Anova(chronhi_foll_c)


#modelling behavior mean durations across high visibility fluctuations
chronhi_chase_d <- update(chronhi_chase_c, duration_s ~ .)
chronhi_lat_d <- update(chronhi_lat_c, duration_s ~ .)
chronhi_foll_d <- update(chronhi_foll_c, duration_s ~ .)

#checking the assumptions
plot(simulateResiduals(fittedModel = chronhi_chase_d)) #assumptions not met
plot(simulateResiduals(fittedModel = chronhi_lat_d)) #assumptions not met
plot(simulateResiduals(fittedModel = chronhi_foll_d))

#transforming the data to meet model assumptions
chronhi_chase_log$duration_s <- log(chronhi_chase_log$duration_s)
chronhi_lat_log$duration_s <- log(chronhi_lat_log$duration_s)

#the transformed models
chrologhi_chase_d <- update(chronhi_chase_d, . ~ ., data = chronhi_chase_log)
chrologhi_lat_d <- update(chronhi_lat_d, . ~ ., data = chronhi_lat_log)

#checking the transformed model assumptions
plot(simulateResiduals(fittedModel = chrologhi_chase_d))
plot(simulateResiduals(fittedModel = chrologhi_lat_d))

#analyzing the results
Anova(chrologhi_chase_d)
Anova(chrologhi_lat_d)
Anova(chronhi_foll_d)

#post hoc pairwise comparison of the significant result
chrologhi_chase_d %>% emmeans(pairwise ~ treatment)


#modelling only the recordings taken when the flux condition had low visibility

#filtering the data by behavior in order to model each behavior separately and
#putting all of them in a list
chronlo_list <- list()
for (i in 1:(length(behavs))){
  chronlo_list[[i]] <- low_chron %>%
                     filter(behavior == behavs[i])
}

#models of behavior counts across low visibility recordings
chronlo_chase_c <- glmmTMB(behav_count ~ treatment*sex + (1|tankID),
                       data = chronlo_list[[1]], family = gaussian())
chronlo_lat_c <- update(chronlo_chase_c, . ~ ., data = chronlo_list[[2]])
chronlo_cont_c <- update(chronlo_chase_c, . ~ ., data = chronlo_list[[3]])
chronlo_foll_c <- update(chronlo_chase_c, . ~ ., data = chronlo_list[[4]])

#checking the model assumptions
plot(simulateResiduals(fittedModel = chronlo_chase_c)) #assumptions not met
plot(simulateResiduals(fittedModel = chronlo_lat_c)) #assumptions not met
plot(simulateResiduals(fittedModel = chronlo_cont_c))
plot(simulateResiduals(fittedModel = chronlo_foll_c)) #assumptions not met

#transforming the data to meet model assumptions
chronlo_chase_log <- as.data.frame(chronlo_list[[1]])
chronlo_chase_log$behav_count <- log(chronlo_chase_log$behav_count+1)
chronlo_lat_log <- as.data.frame(chronlo_list[[2]])
chronlo_lat_log$behav_count <- log(chronlo_lat_log$behav_count+1)
chronlo_foll_log <- as.data.frame(chronlo_list[[4]])
chronlo_foll_log$behav_count <- log(chronlo_foll_log$behav_count+1)

#the transformed models
chrologlo_chase_c <- update(chronlo_chase_c, . ~ ., data = chronlo_chase_log)
chrologlo_lat_c <- update(chronlo_lat_c, . ~ ., data = chronlo_lat_log)
chrologlo_foll_c <- update(chronlo_foll_c, . ~ ., data = chronlo_foll_log)

#checking the transformed model assumptions
plot(simulateResiduals(fittedModel = chrologlo_chase_c))
plot(simulateResiduals(fittedModel = chrologlo_lat_c)) #assumptions not met
plot(simulateResiduals(fittedModel = chrologlo_foll_c))

#analyzing the results
Anova(chrologlo_chase_c)
#Anova(chrologlo_lat_c)
Anova(chronlo_cont_c)
Anova(chrologlo_foll_c)

#post hoc pairwise comparison of the significant result
chrologlo_foll_c %>% emmeans(pairwise ~ treatment)


#modelling  behavior mean durations across low visibility fluctuations
chronlo_chase_d <- update(chronlo_chase_c, duration_s ~ .)
chronlo_lat_d <- update(chronlo_lat_c, duration_s ~ .)
chronlo_foll_d <- update(chronlo_foll_c, duration_s ~ .)

#checking the model assumptions
plot(simulateResiduals(fittedModel = chronlo_chase_d)) #assumptions not met
plot(simulateResiduals(fittedModel = chronlo_lat_d)) #assumptions not met
plot(simulateResiduals(fittedModel = chronlo_foll_d)) #assumptions not met

#transforming the data to meet model assumptions
chronlo_chase_log <- as.data.frame(chronlo_list[[1]])
chronlo_chase_log$duration_s <- log(chronlo_chase_log$duration_s)
chronlo_lat_log$duration_s <- log(chronlo_lat_log$duration_s)
chronlo_foll_log <- as.data.frame(chronlo_list[[4]])
chronlo_foll_log$duration_s <- log(chronlo_foll_log$duration_s)

#the transformed models
chrologlo_chase_d <- update(chronlo_chase_d, . ~ ., data = chronlo_chase_log)
chrologlo_lat_d <- update(chronlo_lat_d, . ~ ., data = chronlo_lat_log)
chrologlo_foll_d <- update(chronlo_foll_d, . ~ ., data = chronlo_foll_log)

#checking the assumptions of the transformed models
plot(simulateResiduals(fittedModel = chrologlo_chase_d))
plot(simulateResiduals(fittedModel = chrologlo_lat_d))
plot(simulateResiduals(fittedModel = chrologlo_foll_d))

#analyzing the results
Anova(chrologlo_chase_d)
Anova(chrologlo_lat_d)
Anova(chrologlo_foll_d)
```


## Direct Comparisons of Count Data

```{r direct count}
#the clay vs dye models

#chases
d_con_cd_chase <- lm(behav_count ~ treatment*experiment, data = direct_claydye[[1]])
#lateral displays
d_con_cd_lat <- update(d_con_cd_chase, . ~ ., data = direct_claydye[[2]])
#aggressive contacts
d_con_cd_cont <- update(d_con_cd_chase, . ~ ., data = direct_claydye[[3]])
#follows
d_con_cd_foll <- update(d_con_cd_cont, . ~ ., data = direct_claydye[[4]])
#aggregated aggressive behaviors
d_con_cd_angr <- update(d_con_cd_chase, . ~ ., data = direct_claydye[[5]])
#all behaviors aggregated
d_con_cd_all <- update(d_con_cd_chase, . ~ ., data = direct_claydye[[6]])

#checking the model assumptions
plot(d_con_cd_chase) #assumptions not met
plot(d_con_cd_lat) #assumptions not met
plot(d_con_cd_cont)
plot(d_con_cd_foll)
plot(d_con_cd_angr)
plot(d_con_cd_all) #assumptions borderline

#transforming the data to meet model assumptions
direct_cd_trans <- direct_claydye
direct_cd_trans[[1]]$behav_count <- sqrt(direct_cd_trans[[1]]$behav_count) +
                                      sqrt(direct_cd_trans[[1]]$behav_count+1)
direct_cd_trans[[2]]$behav_count <- sqrt(direct_cd_trans[[2]]$behav_count) +
                                      sqrt(direct_cd_trans[[2]]$behav_count+1)
direct_cd_trans[[6]]$behav_count <- sqrt(direct_cd_trans[[6]]$behav_count) +
                                      sqrt(direct_cd_trans[[6]]$behav_count+1)

#the transformed models
d_con_cd_chase_trans <- update(d_con_cd_chase, . ~ .,
                               data = direct_cd_trans[[1]])
d_con_cd_lat_trans <- update(d_con_cd_lat, . ~ ., data = direct_cd_trans[[2]])
d_con_cd_all_trans <- update(d_con_cd_all, . ~ ., data = direct_cd_trans[[6]])

#checking the transformed model assumptions
plot(d_con_cd_chase_trans)
plot(d_con_cd_lat_trans)
plot(d_con_cd_all_trans)

#analyzing the results
Anova(d_con_cd_chase_trans)
Anova(d_con_cd_lat_trans)
Anova(d_con_cd_cont)
Anova(d_con_cd_foll)
Anova(d_con_cd_angr)
Anova(d_con_cd_all_trans)


#the dye vs dye models

#chases
d_con_dd_chase <- lm(behav_count ~ treatment*experiment,
                     data = direct_dyedye[[1]])
#lateral displays
d_con_dd_lat <- update(d_con_dd_chase, . ~ ., data = direct_dyedye[[2]])
#aggressive contacts
d_con_dd_cont <- update(d_con_dd_chase, . ~ ., data = direct_dyedye[[3]])
#follows
d_con_dd_foll <- update(d_con_dd_chase, . ~ ., data = direct_dyedye[[4]])
#aggregated aggressive behaviors
d_con_dd_angr <- update(d_con_dd_chase, . ~ ., data = direct_dyedye[[5]])
#all behaviors aggregated
d_con_dd_all <- update(d_con_dd_chase, . ~ ., data = direct_dyedye[[6]])

#checking the model assumptions
plot(d_con_dd_chase) #assumptions borderline
plot(d_con_dd_lat) #qq plot iffy
plot(d_con_dd_cont)
plot(d_con_dd_foll)
plot(d_con_dd_angr)
plot(d_con_dd_all)

#transforming the data to meet model assumptions
direct_dd_trans <- direct_dyedye
direct_dd_trans[[1]]$behav_count <- sqrt(direct_dd_trans[[1]]$behav_count+0.5)
direct_dd_trans[[2]]$behav_count <- sqrt(direct_dd_trans[[2]]$behav_count) +
                                      sqrt(direct_dd_trans[[2]]$behav_count+1)

#the transformed models
d_con_dd_chase_trans <- update(d_con_dd_chase, . ~ .,
                               data = direct_dd_trans[[1]])
d_con_dd_lat_trans <- update(d_con_dd_chase, . ~ .,
                               data = direct_dd_trans[[2]])

#checking the transformed model assumptions
plot(d_con_dd_chase_trans)
plot(d_con_dd_lat_trans)

#analyzing the results
Anova(d_con_dd_chase_trans)
Anova(d_con_dd_lat_trans)
Anova(d_con_dd_cont)
Anova(d_con_dd_foll)
Anova(d_con_dd_angr)
Anova(d_con_dd_all)
```


## Direct Comparison of Duration Data

```{r direct duration}
#the clay vs dye models

#chases
d_dur_cd_chase <- lm(duration_s ~ treatment*experiment, data = direct_claydye[[1]])
#lateral displays
d_dur_cd_lat <- update(d_dur_cd_chase, . ~ ., data = direct_claydye[[2]])
#follows
d_dur_cd_foll <- update(d_dur_cd_chase, . ~ ., data = direct_claydye[[4]])
#aggregated aggressive behaviors
d_dur_cd_angr <- update(d_dur_cd_chase, . ~ ., data = direct_claydye[[5]])
#all behaviors aggregated
d_dur_cd_all <- update(d_dur_cd_chase, . ~ ., data = direct_claydye[[6]])

#checking the model assumptions
plot(d_dur_cd_chase) #assumptions not met
plot(d_dur_cd_lat)
plot(d_dur_cd_foll)
plot(d_dur_cd_angr) #not great
plot(d_dur_cd_all)

#transforming the data to meet model assumptions
direct_cd_trans <- direct_claydye
direct_cd_trans[[1]]$duration_s <- log(direct_cd_trans[[1]]$duration_s)
direct_cd_trans[[5]]$duration_s <- log(direct_cd_trans[[5]]$duration_s)

#the transformed models
d_dur_cd_chase_trans <- update(d_dur_cd_chase, . ~ .,
                               data = direct_cd_trans[[1]])
d_dur_cd_angr_trans <- update(d_dur_cd_angr, . ~ ., data = direct_cd_trans[[5]])

#checking the transformed model assumptions
plot(d_dur_cd_chase_trans)
plot(d_dur_cd_angr_trans)

#analyzing the results
Anova(d_dur_cd_chase_trans)
Anova(d_dur_cd_lat)
Anova(d_dur_cd_foll)
Anova(d_dur_cd_angr_trans)
Anova(d_dur_cd_all)



###NONE OF THE BELOW IS READY
#the dye vs dye models

#chases
d_dur_dd_chase <- lm(duration_s ~ treatment*experiment,
                     data = direct_dyedye[[1]])
#lateral displays
d_dur_dd_lat <- update(d_dur_dd_chase, . ~ ., data = direct_dyedye[[2]])
#follows
d_dur_dd_foll <- update(d_dur_dd_chase, . ~ ., data = direct_dyedye[[4]])
#aggregated aggressive behaviors
d_dur_dd_angr <- update(d_dur_dd_chase, . ~ ., data = direct_dyedye[[5]])
#all behaviors aggregated
d_dur_dd_all <- update(d_dur_dd_chase, . ~ ., data = direct_dyedye[[6]])

#checking the model assumptions
plot(d_dur_dd_chase) #iffy
plot(d_dur_dd_lat)
plot(d_dur_dd_foll)
plot(d_dur_dd_angr)
plot(d_dur_dd_all)

#transforming the data to meet model assumptions
direct_dd_trans <- direct_dyedye
direct_dd_trans[[1]]$duration_s <- log(direct_dd_trans[[1]]$duration_s)

#the transformed model
d_dur_dd_chase_trans <- update(d_dur_dd_chase, . ~ .,
                               data = direct_dd_trans[[1]])

#checking the transformed model assumptions
plot(d_dur_dd_chase_trans)

#analyzing the results
Anova(d_dur_dd_chase_trans)
Anova(d_dur_dd_lat)
Anova(d_dur_dd_foll)
Anova(d_dur_dd_angr)
Anova(d_dur_dd_all)
```


## Analyses for Experiment 2: Acute (Two Day) Clay Exposure
#  Analyses divided by behavior and by high and low visibility fluctuations

```{r acute clay highlow}
#modelling only the recordings taken when the flux condition had high visibility

#filtering the data by behavior in order to model each behavior separately and
#putting all of them in a list
acutehi_list <- list()
for (i in 1:(length(behavs))){
  acutehi_list[[i]] <- high_acute %>%
                     filter(behavior == behavs[i])
}

#models of behavior counts across high visibility fluctuations
acutehi_chase_c <- glmmTMB(behav_count ~ treatment + (1|tankID),
                       data = acutehi_list[[1]], family = gaussian())
acutehi_lat_c <- update(acutehi_chase_c, . ~ ., data = acutehi_list[[2]])
acutehi_cont_c <- update(acutehi_chase_c, . ~ ., data = acutehi_list[[3]])
acutehi_foll_c <- update(acutehi_chase_c, . ~ ., data = acutehi_list[[4]])

#checking the model assumptions
plot(simulateResiduals(fittedModel = acutehi_chase_c))
plot(simulateResiduals(fittedModel = acutehi_lat_c))
plot(simulateResiduals(fittedModel = acutehi_cont_c)) #assumptions not met
plot(simulateResiduals(fittedModel = acutehi_foll_c))

#transforming the data to meet model assumptions
acuhi_cont_log <- as.data.frame(acutehi_list[[3]])
acuhi_cont_log$behav_count <- log(acuhi_cont_log$behav_count+1)

#the transformed models
aculoghi_cont_c <- update(acutehi_cont_c, . ~ ., data = acuhi_cont_log)

#checking the transformed model assumptions
plot(simulateResiduals(fittedModel = aculoghi_cont_c))

#analyzing the models
Anova(acutehi_chase_c)
Anova(acutehi_lat_c)
Anova(aculoghi_cont_c)
Anova(acutehi_foll_c)


#modelling  behavior mean durations across high visibility fluctuations
acutehi_chase_d <- update(acutehi_chase_c, duration_s ~ .)
acutehi_lat_d <- update(acutehi_lat_c, duration_s ~ .)
acutehi_foll_d <- update(acutehi_foll_c, duration_s ~ .)

#checking the model assumptions
plot(simulateResiduals(fittedModel = acutehi_chase_d))
plot(simulateResiduals(fittedModel = acutehi_lat_d))
plot(simulateResiduals(fittedModel = acutehi_foll_d)) #doesn't meet model assumptions

#transforming the data to meet model assumptions
acuhi_foll_log <- as.data.frame(acutehi_list[[4]])
acuhi_foll_log$duration_s <- log(acuhi_foll_log$duration_s)

#the transformed model
aculoghi_foll_d <- update(acutehi_foll_d, . ~ ., data = acuhi_foll_log)

#checking the transformed model assumptions
plot(simulateResiduals(aculoghi_foll_d))

#analyzing the models
Anova(acutehi_chase_d)
Anova(acutehi_lat_d)
Anova(aculoghi_foll_d)

#post hoc pairwise comparison of the significant results
acutehi_lat_d %>% emmeans(pairwise ~ treatment)


#modelling only the recordings taken when the flux condition had low visibility

#filtering the data by behavior in order to model each behavior separately and
#putting all of them in a list
acutelo_list <- list()
for (i in 1:(length(behavs))){
  acutelo_list[[i]] <- low_acute %>%
                     filter(behavior == behavs[i])
}

#models of behavior counts across low visibility recordings
##random effect excluded from chase count model due to singular fit
acutelo_chase_c <- lm(behav_count ~ treatment, data = acutelo_list[[1]])
acutelo_lat_c <- glmmTMB(behav_count ~ treatment + (1|tankID),
                         data = acutelo_list[[2]], family = gaussian)
acutelo_cont_c <- update(acutelo_lat_c, . ~ ., data = acutelo_list[[3]])
acutelo_foll_c <- update(acutelo_lat_c, . ~ ., data = acutelo_list[[4]])

#checking the model assumptions
plot(simulateResiduals(fittedModel = acutelo_chase_c))
plot(simulateResiduals(fittedModel = acutelo_lat_c))
plot(simulateResiduals(fittedModel = acutelo_cont_c))
plot(simulateResiduals(fittedModel = acutelo_foll_c))

#analyzing the results
Anova(acutelo_chase_c)
Anova(acutelo_lat_c)
Anova(acutelo_cont_c)
Anova(acutelo_foll_c)


#modelling  behavior mean durations across high visibility fluctuations
##random effect excluded from chase model due to singular fit
acutelo_chase_d <- update(acutelo_chase_c, duration_s ~ .)
acutelo_lat_d <- update(acutelo_lat_c, duration_s ~ .)
acutelo_foll_d <- update(acutelo_foll_c, duration_s ~ .)

#checking the model assumptions
plot(simulateResiduals(fittedModel = acutelo_chase_d))
plot(simulateResiduals(fittedModel = acutelo_lat_d))
plot(simulateResiduals(fittedModel = acutelo_foll_d))

#analyzing the results
Anova(acutelo_chase_d)
Anova(acutelo_lat_d)
Anova(acutelo_foll_d)
```


## Analyses for Experiment 3: Stirred (Two Day) Dye Exposure
#  Analyses divided by behavior and including all recordings

```{r dye stirred}
#filtering the data by behavior in order to model each behavior separately and
#putting all of them in a list
stir_list <- list()
for (i in 1:(length(behavs))){
  stir_list[[i]] <- stir_dye %>%
                     filter(behavior == behavs[i])
}

#models of behavior counts across all recordings
stir_chase_c <- glmmTMB(behav_count ~ treatment + (1|tankID),
                       data = stir_list[[1]], family = gaussian())
stir_lat_c <- update(stir_chase_c, . ~ ., data = stir_list[[2]])
stir_cont_c <- update(stir_chase_c, . ~ ., data = stir_list[[3]])
stir_foll_c <- update(stir_chase_c, . ~ ., data = stir_list[[4]])

#checking the model assumptions
plot(simulateResiduals(fittedModel = stir_chase_c))
plot(simulateResiduals(fittedModel = stir_lat_c)) #assumptions not met
plot(simulateResiduals(fittedModel = stir_cont_c)) #assumptions not met
plot(simulateResiduals(fittedModel = stir_foll_c)) #assumptions not met

#transforming the data to meet model assumptions
stir_lat_log <- as.data.frame(stir_list[[2]])
stir_lat_log$behav_count <- log(stir_lat_log$behav_count+1)
stir_cont_log <- as.data.frame(stir_list[[3]])
stir_cont_log$behav_count <- log(stir_cont_log$behav_count+1)
stir_foll_log <- as.data.frame(stir_list[[4]])
stir_foll_log$behav_count <- log(stir_foll_log$behav_count+1)

#the transformed models
stirlog_lat_c <- update(stir_lat_c, . ~ ., data = stir_lat_log)
stirlog_cont_c <- update(stir_cont_c, . ~ ., data = stir_cont_log)
stirlog_foll_c <- update(stir_foll_c, . ~ ., data = stir_foll_log)

#testing the transformed model assumptions
plot(simulateResiduals(fittedModel = stirlog_lat_c))
plot(simulateResiduals(fittedModel = stirlog_cont_c))
plot(simulateResiduals(fittedModel = stirlog_foll_c))

#analyzing the results
Anova(stir_chase_c)
Anova(stirlog_lat_c)
Anova(stirlog_cont_c)
Anova(stirlog_foll_c)


#models of behavior durations across all relevant recordings
stir_chase_d <- update(stir_chase_c, duration_s ~ .)
stir_lat_d <- update(stir_lat_c, duration_s ~ .)
stir_foll_d <- update(stir_foll_c, duration_s ~ .)

#checking the assumptions
plot(simulateResiduals(fittedModel = stir_chase_d)) #assumptions not met
plot(simulateResiduals(fittedModel = stir_lat_d)) #assumptions not met
plot(simulateResiduals(fittedModel = stir_foll_d)) #assumptions not met

#transforming the data to meet model assumptions
stir_chase_log <- as.data.frame(stir_list[[1]])
stir_chase_log$duration_s <- log(stir_chase_log$duration_s)
stir_lat_log$duration_s <- log(stir_lat_log$duration_s)
stir_foll_log$duration_s <- log(stir_foll_log$duration_s)

#the transformed models
stirlog_chase_d <- update(stir_chase_d, . ~ ., data = stir_chase_log)
stirlog_lat_d <- update(stir_lat_d, . ~ ., data = stir_lat_log)
stirlog_foll_d <- update(stir_foll_d, . ~ ., data = stir_foll_log)

#testing the assumptions of the transformed models
plot(simulateResiduals(fittedModel = stirlog_chase_d))
plot(simulateResiduals(fittedModel = stirlog_lat_d))
plot(simulateResiduals(fittedModel = stirlog_foll_d))

#analyzing the results
Anova(stirlog_chase_d)
Anova(stirlog_lat_d)
Anova(stirlog_foll_d)
```


## Analyses for Experiment 3: Stirred (Two Day) Dye Exposure
#  Analyses divided by behavior and by high and low visibility fluctuations

```{r dye stirred highlow}
#modelling only the recordings taken when the flux condition had high visibility

#filtering the data by behavior in order to model each behavior separately and
#putting all of them in a list
stirhi_list <- list()
for (i in 1:(length(behavs))){
  stirhi_list[[i]] <- high_stir %>%
                     filter(behavior == behavs[i])
}

#models of behavior counts across high visibility fluctuations
stirhi_chase_c <- glmmTMB(behav_count ~ treatment + (1|tankID),
                       data = stirhi_list[[1]], family = gaussian)
stirhi_lat_c <- update(stirhi_chase_c, . ~ ., data = stirhi_list[[2]])
stirhi_cont_c <- update(stirhi_chase_c, . ~ ., data = stirhi_list[[3]])
stirhi_foll_c <- update(stirhi_chase_c, . ~ ., data = stirhi_list[[4]])

#checking the model assumptions
plot(simulateResiduals(fittedModel = stirhi_chase_c))
plot(simulateResiduals(fittedModel = stirhi_lat_c)) #assumptions not met
plot(simulateResiduals(fittedModel = stir_cont_c)) #assumptions not met
plot(simulateResiduals(fittedModel = stirhi_foll_c))

#transforming the data to meet model assumptions
stirhi_lat_log <- as.data.frame(stirhi_list[[2]])
stirhi_lat_log$behav_count <- log(stirhi_lat_log$behav_count+1)
stirhi_cont_log <- as.data.frame(stirhi_list[[3]])
stirhi_cont_log$behav_count <- log(stirhi_cont_log$behav_count+1)

#the transformed models
stirloghi_lat_c <- update(stirhi_lat_c, . ~ ., data = stirhi_lat_log)
stirloghi_cont_c <- update(stirhi_cont_c, . ~ ., data = stirhi_cont_log)

#checking the transformed model assumptions
plot(simulateResiduals(fittedModel = stirloghi_lat_c))
plot(simulateResiduals(fittedModel = stirloghi_cont_c))

#analyzing the results
Anova(stirhi_chase_c)
Anova(stirloghi_lat_c)
Anova(stirloghi_cont_c)
Anova(stirhi_foll_c)


#modellingbehavior mean durations across high visibility fluctuations
stirhi_chase_d <- update(stirhi_chase_c, duration_s ~ .)
stirhi_lat_d <- update(stirhi_lat_c, duration_s ~ .)
stirhi_foll_d <- update(stirhi_foll_c, duration_s ~ .)

#checking the model assumptions
plot(simulateResiduals(fittedModel = stirhi_chase_d)) #assumptions not met
plot(simulateResiduals(fittedModel = stirhi_lat_d)) #assumptions not met
plot(simulateResiduals(fittedModel = stirhi_foll_d))

#transforming the data to meet model assumptions
stirhi_chase_log <- as.data.frame(stirhi_list[[1]])
stirhi_chase_log$duration_s <- log(stirhi_chase_log$duration_s)
stirhi_lat_log <- as.data.frame(stirhi_list[[2]])
stirhi_lat_log$duration_s <- log(stirhi_lat_log$duration_s)

#the transformed models
stirloghi_chase_d <- update(stirhi_chase_d, . ~ ., data = stirhi_chase_log)
stirloghi_lat_d <- update(stirhi_lat_d, . ~ ., data = stirhi_lat_log)

#checking the assumptions of the transformed models
plot(simulateResiduals(fittedModel = stirloghi_chase_d))
plot(simulateResiduals(fittedModel = stirloghi_lat_d))

#analyzing the results
Anova(stirloghi_chase_d)
Anova(stirloghi_lat_d)
Anova(stirhi_foll_d)


#modelling only the recordings taken when the flux condition had low visibility

#filtering the data by behavior in order to model each behavior separately and
#putting all of them in a list
stirlo_list <- list()
for (i in 1:(length(behavs))){
  stirlo_list[[i]] <- low_stir %>%
                     filter(behavior == behavs[i])
}

#models of behavior counts across low visibility recordings
stirlo_chase_c <- glmmTMB(behav_count ~ treatment + (1|tankID),
                       data = stirlo_list[[1]], family = gaussian)
stirlo_lat_c <- update(stirlo_chase_c, . ~ ., data = stirlo_list[[2]])
stirlo_cont_c <- update(stirlo_chase_c, . ~ ., data = stirlo_list[[3]])
stirlo_foll_c <- update(stirlo_chase_c, . ~ ., data = stirlo_list[[4]])

#checking the model assumptions
plot(simulateResiduals(fittedModel = stirlo_chase_c)) #assumptions not met
plot(simulateResiduals(fittedModel = stirlo_lat_c)) #assumptions not met
plot(simulateResiduals(fittedModel = stirlo_cont_c))
plot(simulateResiduals(fittedModel = stirlo_foll_c)) #assumptions not met

#transforming the data to meet model assumptions
stirlo_chase_log <- as.data.frame(stirlo_list[[1]])
stirlo_chase_log$behav_count <- log(stirlo_chase_log$behav_count+1)
stirlo_lat_log <- as.data.frame(stirlo_list[[2]])
stirlo_lat_log$behav_count <- log(stirlo_lat_log$behav_count+1)
stirlo_foll_log <- as.data.frame(stirlo_list[[4]])
stirlo_foll_log$behav_count <- log(stirlo_foll_log$behav_count+1)

#the transformed models
stirloglo_chase_c <- update(stirlo_chase_c, . ~ ., data = stirlo_chase_log)
stirloglo_lat_c <- update(stirlo_lat_c, . ~ ., data = stirlo_lat_log)
stirloglo_foll_c <- update(stirlo_foll_c, . ~ ., data = stirlo_foll_log)

#checking the model assumptions
plot(simulateResiduals(fittedModel = stirloglo_chase_c))
plot(simulateResiduals(fittedModel = stirloglo_lat_c))
plot(simulateResiduals(fittedModel = stirloglo_foll_c))

#analyzing the results
Anova(stirloglo_chase_c)
Anova(stirloglo_lat_c)
Anova(stirlo_cont_c)
Anova(stirloglo_foll_c)


#modelling  behavior mean durations across low visibility fluctuations
stirlo_chase_d <- update(stirlo_chase_c, duration_s ~ .)
stirlo_lat_d <- update(stirlo_lat_c, duration_s ~ .)
stirlo_foll_d <- update(stirlo_foll_c, duration_s ~ .)

#checking the model assumptions
plot(simulateResiduals(fittedModel = stirlo_chase_d))
plot(simulateResiduals(fittedModel = stirlo_lat_d))
plot(simulateResiduals(fittedModel = stirlo_foll_d))

#analyzing the results
Anova(stirlo_chase_d)
Anova(stirlo_lat_d)
Anova(stirlo_foll_d)
```


## Analyses for Experiment 4: Unstirred (Two Day) Dye Exposure
#  Analyses divided by behavior and including all recordings

```{r dye nostir}
#filtering the data by behavior in order to model each behavior separately and
#putting all of them in a list
nostir_list <- list()
for (i in 1:(length(behavs))){
  nostir_list[[i]] <- nostir_dye %>%
                     filter(behavior == behavs[i])
}

#models of behavior counts across all recordings
nostir_chase_c <- glmmTMB(behav_count ~ treatment + (1|tankID),
                       data = nostir_list[[1]], family = gaussian())
nostir_lat_c <- update(nostir_chase_c, . ~ ., data = nostir_list[[2]])
nostir_cont_c <- update(nostir_chase_c, . ~ ., data = nostir_list[[3]])
nostir_foll_c <- update(nostir_chase_c, . ~ ., data = nostir_list[[4]])

#checking the model assumptions
plot(simulateResiduals(fittedModel = nostir_chase_c)) #assumptions not met
plot(simulateResiduals(fittedModel = nostir_lat_c)) #assumptions not met
plot(simulateResiduals(fittedModel = nostir_cont_c)) #assumptions not met
plot(simulateResiduals(fittedModel = nostir_foll_c))

#transforming the data to meet model assumptions
nostir_chase_log <- as.data.frame(nostir_list[[1]])
nostir_chase_log$behav_count <- log(nostir_chase_log$behav_count+1)
nostir_lat_log <- as.data.frame(nostir_list[[2]])
nostir_lat_log$behav_count <- log(nostir_lat_log$behav_count+1)
nostir_cont_log <- as.data.frame(nostir_list[[3]])
nostir_cont_log$behav_count <- log(nostir_cont_log$behav_count+1)

#the transformed models
nolog_chase_c <- update(nostir_chase_c, . ~ ., data = nostir_chase_log)
nolog_lat_c <- update(nostir_lat_c, . ~ ., data = nostir_lat_log)
nolog_cont_c <- update(nostir_cont_c, . ~ ., data = nostir_cont_log)

#checking the transformed model assumptions
plot(simulateResiduals(fittedModel = nolog_chase_c))
plot(simulateResiduals(fittedModel = nolog_lat_c))
plot(simulateResiduals(fittedModel = nolog_cont_c))

#analyzing the results
Anova(nolog_chase_c)
Anova(nolog_lat_c)
Anova(nolog_cont_c)
Anova(nostir_foll_c)


#models of behavior durations across all relevant recordings
nostir_chase_d <- update(nostir_chase_c, duration_s ~ .)
nostir_lat_d <- update(nostir_lat_c, duration_s ~ .)
nostir_foll_d <- update(nostir_foll_c, duration_s ~ .)

#checking the model assumptions
plot(simulateResiduals(fittedModel = nostir_chase_d)) #assumptions not met
plot(simulateResiduals(fittedModel = nostir_lat_d)) #assumptions not met
plot(simulateResiduals(fittedModel = nostir_foll_d))

#transforming the data to meet model assumptions
nostir_chase_log$duration_s <- log(nostir_chase_log$duration_s)
nostir_lat_log$duration_s <- log(nostir_lat_log$duration_s)

#the transformed models
nolog_chase_d <- update(nostir_chase_d, . ~ ., data = nostir_chase_log)
nolog_lat_d <- update(nostir_lat_d, . ~ ., data = nostir_lat_log)

#checking the transformed model assumptions
plot(simulateResiduals(fittedModel = nolog_chase_d))
plot(simulateResiduals(fittedModel = nolog_lat_d))

#analyzing the results
Anova(nolog_chase_d)
Anova(nolog_lat_d)
Anova(nostir_foll_d)
```


## Analyses for Experiment 4: Unstirred (Two Day) Dye Exposure
#  Analyses divided by behavior and by high and low visibility fluctuations

```{r dye nostir highlow}
#modelling only the recordings taken when the flux condition had high visibility

#filtering the data by behavior in order to model each behavior separately and
#putting all of them in a list
nostirhi_list <- list()
for (i in 1:(length(behavs))){
  nostirhi_list[[i]] <- high_nostir %>%
                     filter(behavior == behavs[i])
}

#models of behavior counts across high visibility recordings
nostirhi_chase_c <- glmmTMB(behav_count ~ treatment + (1|tankID),
                       data = nostirhi_list[[1]], family = gaussian)
nostirhi_lat_c <- update(nostirhi_chase_c, . ~ ., data = nostirhi_list[[2]])
nostirhi_cont_c <- update(nostirhi_chase_c, . ~ ., data = nostirhi_list[[3]])
nostirhi_foll_c <- update(nostirhi_chase_c, . ~ ., data = nostirhi_list[[4]])

#checking the model assumptions
plot(simulateResiduals(fittedModel = nostirhi_chase_c)) #assumptions not met
plot(simulateResiduals(fittedModel = nostirhi_lat_c)) #assumptions not met
plot(simulateResiduals(fittedModel = nostirhi_cont_c))
plot(simulateResiduals(fittedModel = nostirhi_foll_c))

#transforming the data to meet model assumptions
nostirhi_chase_log <- as.data.frame(nostirhi_list[[1]])
nostirhi_chase_log$behav_count <- log(nostirhi_chase_log$behav_count+1)
nostirhi_lat_log <- as.data.frame(nostirhi_list[[2]])
nostirhi_lat_log$behav_count <- log(nostirhi_lat_log$behav_count+1)

#the transformed models
nologhi_chase_c <- update(nostirhi_chase_c, . ~ ., data = nostirhi_chase_log)
nologhi_lat_c <- update(nostirhi_lat_c, . ~ ., data = nostirhi_lat_log)

#checking the assumptions of the transformed models
plot(simulateResiduals(fittedModel = nologhi_chase_c))
plot(simulateResiduals(fittedModel = nologhi_lat_c))

#analyzing the results
Anova(nologhi_chase_c)
Anova(nologhi_lat_c)
Anova(nostirhi_cont_c)
Anova(nostirhi_foll_c)

#post hoc pairwise comparison of the significant result
nologhi_lat_c %>% emmeans(pairwise ~ treatment)
nostirhi_foll_c %>% emmeans(pairwise ~ treatment)


#modelling  behavior mean durations across high visibility fluctuations
nostirhi_chase_d <- update(nostirhi_chase_c, duration_s ~ .)
nostirhi_lat_d <- update(nostirhi_lat_c, duration_s ~ .)
nostirhi_foll_d <- update(nostirhi_foll_c, duration_s ~ .)

#checking the model assumptions
plot(simulateResiduals(fittedModel = nostirhi_chase_d)) #assumptions not met
plot(simulateResiduals(fittedModel = nostirhi_lat_d)) #assumptions not met
plot(simulateResiduals(fittedModel = nostirhi_foll_d))

#transforming the data to meet model assumptions
nostirhi_chase_log$duration_s <- log(nostirhi_chase_log$duration_s)
nostirhi_lat_log$duration_s <- log(nostirhi_lat_log$duration_s)

#the transformed models
nologhi_chase_d <- update(nostirhi_chase_d, . ~ ., data = nostirhi_chase_log)
nologhi_lat_d <- update(nostirhi_lat_d, . ~ ., data = nostirhi_lat_log)

#checking the transformed model assumptions
plot(simulateResiduals(fittedModel = nologhi_chase_d))
plot(simulateResiduals(fittedModel = nologhi_lat_d))

#analyzing the results
Anova(nologhi_chase_d)
Anova(nologhi_lat_d)
Anova(nostirhi_foll_d)


#modelling only the recordings taken when the flux condition had low visibility

#filtering the data by behavior in order to model each behavior separately and
#putting all of them in a list
nostirlo_list <- list()
for (i in 1:(length(behavs))){
  nostirlo_list[[i]] <- low_nostir %>%
                     filter(behavior == behavs[i])
}

#models of behavior counts across low visibility recordings
nostirlo_chase_c <- glmmTMB(behav_count ~ treatment + (1|tankID),
                       data = nostirlo_list[[1]])
nostirlo_lat_c <- update(nostirlo_chase_c, . ~ ., data = nostirlo_list[[2]])
nostirlo_cont_c <- update(nostirlo_chase_c, . ~ ., data = nostirlo_list[[3]])
nostirlo_foll_c <- update(nostirlo_chase_c, . ~ ., data = nostirlo_list[[4]])

#checking the model assumptions
plot(simulateResiduals(fittedModel = nostirlo_chase_c))
plot(simulateResiduals(fittedModel = nostirlo_lat_c)) #assumptions not met
plot(simulateResiduals(fittedModel = nostirlo_cont_c)) #assumptions not met
plot(simulateResiduals(fittedModel = nostirlo_foll_c))

#transforming the data to meet model assumptions
nostirlo_lat_log <- as.data.frame(nostirlo_list[[2]])
nostirlo_lat_log$behav_count <- log(nostirlo_lat_log$behav_count+1)
nostirlo_cont_log <- as.data.frame(nostirlo_list[[3]])
nostirlo_cont_log$behav_count <- log(nostirlo_cont_log$behav_count+1)

#the transformed models
nologlo_lat_c <- update(nostirlo_lat_c, . ~ ., data = nostirlo_lat_log)
nologlo_cont_c <- update(nostirlo_cont_c, . ~ ., data = nostirlo_cont_log)

#checking the transformed model assumptions
plot(simulateResiduals(fittedModel = nologlo_lat_c)) #assumptions not met
plot(simulateResiduals(fittedModel = nologlo_cont_c))

#alternative transformation for the one that doesn't work
nostirlo_lat_recip <- as.data.frame(nostirlo_list[[2]])
nostirlo_lat_recip$behav_count <- (1/(nostirlo_lat_recip$behav_count+1))
noreciplo_lat_c <- update(nostirlo_lat_c, . ~ ., data = nostirlo_lat_recip)
plot(simulateResiduals(fittedModel = noreciplo_lat_c))

#analyzing the results
Anova(nostirlo_chase_c)
Anova(noreciplo_lat_c)
Anova(nologlo_cont_c)
Anova(nostirlo_foll_c)


#modellingbehavior mean durations across low visibility fluctuations only
nostirlo_chase_d <- update(nostirlo_chase_c, duration_s ~ .)
nostirlo_lat_d <- update(nostirlo_lat_c, duration_s ~ .)
nostirlo_foll_d <- update(nostirlo_foll_c, duration_s ~ .)

#checking the model assumptions
plot(simulateResiduals(fittedModel = nostirlo_chase_d))
plot(simulateResiduals(fittedModel = nostirlo_lat_d))
plot(simulateResiduals(fittedModel = nostirlo_foll_d))

#analyzing the models
Anova(nostirlo_chase_d)
Anova(nostirlo_lat_d)
Anova(nostirlo_foll_d)

#post hoc pairwise comparison of the significant result
nostirlo_lat_d %>% emmeans(pairwise ~ treatment)
```
