---
title: "BORIS Data Tidying"
author: "Hannah Anderson"
date: "2023-04-12"
output: pdf_document
---

## Setup

Our scorers used BORIS to score the behavior of fish in videos from four
different experiments: a chronic clay exposure, an acute clay exposure, a pond
dye with stirring exposure, and a pond dye without stirring exposure.

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)

#the data files produced by our video scorers
longclay_isabella <- read_csv("chronic_clay_BORIS_Isabella.csv")
longclay_izma <- read_csv("chronic_clay_BORIS_Izma_HMAedited.csv")
shortclay_grace <- read_csv("acute_clay_BORIS_Grace.csv")
shortclay_sana <- read_csv("acute_clay_BORIS_Sana.csv")
dyestir_ryan <- read_csv("stir_dye_BORIS_Ryan_HMAedited.csv")
dye_grace <- read_csv("dye_BORIS_Grace.csv") #includes both dye experiments
nostir_ryan <- read_csv("nostir_dye_BORIS_Ryan_HMAedited.csv")
```


## Cleaning Base Data

Fixing any mistakes or inconsistencies in naming, any Excel date conversion
problems and combining the data from the same experiments together.

```{r tidying and combining}
#Grace included two experiments in her dye file, so I'm separating them
dyestir_grace <- dplyr::filter(dye_grace, grepl("YES", `Media file`))
nostir_grace <- dplyr::filter(dye_grace, grepl("NO", `Media file`))

#trimming down the data frames so they only contain useful columns
longclay_isabella <- longclay_isabella[,c(6:9,12)]
longclay_izma <- longclay_izma[,c(7:8,11,14,17)]
shortclay_grace <- shortclay_grace[,c(7:8,10,13,16)]
shortclay_sana <- shortclay_sana[,c(7:8,10,13,16)]
dyestir_ryan <- dyestir_ryan[,c(7:8,11,14,17)]
dyestir_grace <- dyestir_grace[,c(7:8,10,13,16)]
nostir_ryan <- nostir_ryan[,c(7:8,11,14,17)]
nostir_grace <- nostir_grace[,c(7:8,10,13,16)]

#changing column names to be universal for all experiments
sharecol <- c("tankID", "recording", "behavior", "behav_type", "duration_s")
colnames(longclay_isabella) <- sharecol
colnames(longclay_izma) <- sharecol
colnames(shortclay_grace) <- sharecol
colnames(shortclay_sana) <- sharecol
colnames(dyestir_grace) <- sharecol
colnames(dyestir_ryan) <- sharecol
colnames(nostir_grace) <- sharecol
colnames(nostir_ryan) <- sharecol

#Ryan's stirred dye file is missing the tankID and recording number for the
#tank 7-3, recording 3 video
dyestir_ryan$tankID[is.na(dyestir_ryan$tankID)] <- "7-3"
dyestir_ryan$recording[is.na(dyestir_ryan$recording)] <- 3
#Ryan also used two different behavior names than the others, consolidating
dyestir_ryan$behavior[dyestir_ryan$behavior == "chase"] <- "chasing"
dyestir_ryan$behavior[dyestir_ryan$behavior == "agg"] <- "agg_contact"
nostir_ryan$behavior[nostir_ryan$behavior == "Chasing"] <- "chasing"
nostir_ryan$behavior[nostir_ryan$behavior == "aggressive contact"] <- "agg_contact"
nostir_ryan$behavior[nostir_ryan$behavior == "lateral display"] <- "lat_display"


#Isabella's tank IDs got converted to dates, so converting them back here
datlength <- length(longclay_isabella$tankID)
for (i in 1:datlength) {
  str_hold <- strsplit(longclay_isabella$tankID[i], split = "/")
  longclay_isabella$tankID[i] <- paste(str_hold[[1]][1], str_hold[[1]][2],
                                       sep = "-")
}

#combining the files from different scorers for their shared experiments
all_data <- list()
#a reference to keep track of which item in the list is which, serves no other
#purpose
list_order_ref <- c("1: long clay", "2: short clay", "3: stirred dye",
                    "4: unstirred dye")
all_data[[1]] <- rbind(longclay_isabella, longclay_izma)
all_data[[2]] <- rbind(shortclay_grace, shortclay_sana)
all_data[[3]] <- rbind(dyestir_grace, dyestir_ryan)
all_data[[4]] <- rbind(nostir_grace, nostir_ryan)

#adding a treatment condition column to the files missing it and also sex to the
#long clay experiment
##long clay experiment
all_data[[1]]$treatment <- all_data[[1]]$tankID %>%
                           fct_collapse(flux = c("5-3", "1-1", "4-3", "6-2",
                                                     "3-2", "5-1", "2-3", "8-3"),
                                            stable = c("2-1", "3-1", "6-1", "7-1",
                                                       "4-2", "2-2", "6-3", "7-2"),
                                            clear = c("7-3", "8-2", "8-1", "4-1",
                                                      "3-3", "1-3", "5-2", "1-2"))
#long clay experiment
all_data[[1]]$sex <- all_data[[1]]$tankID %>%
                     fct_collapse(female = c("5-3", "1-1", "4-3", "6-2",
                                             "2-1", "3-1", "6-1", "7-1",
                                             "7-3", "8-2", "8-1", "4-1"),
                                  male = c("3-2", "5-1", "2-3", "8-3",
                                           "4-2", "2-2", "6-3", "7-2",
                                           "3-3", "1-3", "5-2", "1-2"))
#short clay experiment
all_data[[2]]$treatment <- all_data[[2]]$tankID %>%
                           fct_collapse(flux = c("7-3", "3-1", "6-1", "6-2",
                                                 "4-1", "2-1", "7-2", "8-2"),
                                        stable = c("7-1", "3-3", "8-3", "3-2",
                                                   "6-3", "5-1", "1-3", "4-2"),
                                        clear = c("5-3", "1-2", "4-3", "2-2",
                                                  "1-1", "8-1", "2-3", "5-2"))
#dye with stirring experiment
all_data[[3]]$treatment <- all_data[[3]]$tankID %>%
                           fct_collapse(flux = c("1-2", "3-3", "2-1", "4-3",
                                                 "5-1", "6-1", "5-2", "3-1"),
                                        stable = c("4-2", "7-1", "6-3", "2-2",
                                                   "5-3", "8-3", "8-1", "2-3"),
                                        clear = c("7-2", "8-2", "7-3", "4-1",
                                                  "3-2", "1-3", "1-1", "6-2"))
#dye without stirring experiment
all_data[[4]]$treatment <- all_data[[4]]$tankID %>%
                           fct_collapse(flux = c("8-2", "5-1", "3-2", "1-1",
                                                 "6-1", "4-3", "6-3", "1-2"),
                                        stable = c("3-1", "7-1", "3-3", "8-1",
                                                   "2-3", "4-2", "4-1", "8-3"),
                                        clear = c("2-2", "2-1", "5-2", "7-3",
                                                  "7-2", "5-3", "1-3", "6-2"))
```


## Summarizing Data

Summarizing and collapsing the data together so it's in a more usable form for
analysis.

``` {r collapse}
#collapsing the behaviors into counts and durations
collapse_data <- list()
#collapsing the two week clay experiment separately because it includes the
#variable "sex," which the other experiments do not
collapse_data[[1]] <- all_data[[1]] %>%
                      group_by(tankID, recording, behavior, treatment, sex) %>%
                      summarize(duration_s = mean(duration_s),
                                behav_count = as.numeric(table(behavior)))
#collapsing the three remaining experiments
for (i in 2:4) {
  collapse_data[[i]] <- all_data[[i]] %>%
                      group_by(tankID, recording, behavior, treatment) %>%
                      summarize(duration_s = mean(duration_s),
                                behav_count = as.numeric(table(behavior)))
}

#new goal: leave the collapse_data the way it is and create a separate dataframe
#with the new values
temp_data <- list()

#aggregating all of the durations of all behaviors per recording to investigate
#overall behavioral duration trends
##aggregating for the two week clay experiment
temp_data[[1]] <- aggregate(duration_s ~ tankID + recording + treatment + sex,
                            data = all_data[[1]], FUN = "mean")
#aggregating the other three experiments
for (i in 2:4) {
  temp_data[[i]] <- aggregate(duration_s ~ tankID + recording + treatment,
                            data = all_data[[i]], FUN = "mean")
}
#identifying that these values include all behaviors
for (i in 1:4) {
  temp_data[[i]]$behavior <- "all"
}
#moving the data out of the temporary list
behav_all <- temp_data

#aggregating all of the counts of all the behaviors per recording to investigate
#overall behavioral occurence trends
##clearing temp_data to avoid any mishaps
temp_data <- list()
#aggregating the two week experiment
temp_data[[1]] <- aggregate(behav_count ~ tankID + recording + treatment + sex,
                            data = collapse_data[[1]], FUN = "sum")
#aggregating the other three experiments
for (i in 2:4) {
  temp_data[[i]] <- aggregate(behav_count ~ tankID + recording + treatment,
                              data = collapse_data[[i]], FUN = "sum")
}
#identifying that these values include all behaviors
for (i in 1:4) {
  temp_data[[i]]$behavior <- "all"
}
#adding the count data to the duration data for the first experiment
behav_all[[1]] <- full_join(behav_all[[1]], temp_data[[1]], by = c("tankID",
                                                                   "recording",
                                                                   "treatment",
                                                                   "behavior",
                                                                   "sex"))
#adding the count data to the duration data for the other three experiments
for (i in 2:4) {
  behav_all[[i]] <- full_join(behav_all[[i]], temp_data[[i]], by = c("tankID",
                                                                     "recording",
                                                                     "treatment",
                                                                     "behavior"))
}

#aggregating the durations of just aggressive behaviors for each recording to
#investigate the overall trends in aggressive durations
#aggregating all of the durations of all behaviors per recording to investigate
#overall behavioral duration trends
##clearing temp_data to avoid any mishaps
temp_data <- list()
#aggregating for the two week clay experiment
temp_data[[1]] <- all_data[[1]] %>%
                  filter(behavior %in% c("chasing", "lat_display",
                                         "agg_contact")) %>%
                  aggregate(duration_s ~ tankID + recording + treatment + sex,
                            FUN = "mean")
#aggregating the other three experiments
for (i in 2:4) {
  temp_data[[i]] <- all_data[[i]] %>%
                    filter(behavior %in% c("chasing", "lat_display",
                                           "agg_contact")) %>%
                    aggregate(duration_s ~ tankID + recording + treatment,
                              FUN = "mean")
}
#identifying that these values include all aggressive behaviors
for (i in 1:4) {
  temp_data[[i]]$behavior <- "all_aggressive"
}
#moving the data out of the temporary list
behav_agg <- temp_data

#aggregating all of the counts of just aggressive behaviors for each recording
#to investigate trends in overall aggressive occurrences
##clearing temp_data to avoid any mishaps
temp_data <- list()
#aggregating the two week experiment
temp_data[[1]] <- collapse_data[[1]] %>%
                  filter(behavior %in% c("chasing", "lat_display",
                                         "agg_contact")) %>%
                  aggregate(behav_count ~ tankID + recording + treatment + sex,
                            FUN = "sum")
#aggregating the other three experiments
for (i in 2:4) {
  temp_data[[i]] <- collapse_data[[i]] %>%
                    filter(behavior %in% c("chasing", "lat_display",
                                           "agg_contact")) %>%
                    aggregate(behav_count ~ tankID + recording + treatment,
                              FUN = "sum")
}
#identifying that these values include all aggressive behaviors
for (i in 1:4) {
  temp_data[[i]]$behavior <- "all_aggressive"
}
#adding the count data to the duration data for the first experiment
behav_agg[[1]] <- full_join(behav_agg[[1]], temp_data[[1]], by = c("tankID",
                                                                   "recording",
                                                                   "treatment",
                                                                   "behavior",
                                                                   "sex"))
#adding the count data to the duration data for the other three experiments
for (i in 2:4) {
  behav_agg[[i]] <- full_join(behav_agg[[i]], temp_data[[i]], by = c("tankID",
                                                                     "recording",
                                                                     "treatment",
                                                                     "behavior"))
}
#combining the aggression-only list and the all-behavior list for the first
#experiment
behav_all[[1]] <- rbind(behav_all[[1]], behav_agg[[1]])
#and combining them for the three remaining experiments
for (i in 2:4) {
  behav_all[[i]] <- rbind(behav_all[[i]], behav_agg[[i]])
}
```


## Writing Files

Writing the final converted data frames to .csv files

``` {r write}
#for the data separated by individual behaviors
write_csv(collapse_data[[1]], "chronic_clay_BORIS_all.csv")
write_csv(collapse_data[[2]], "acute_clay_BORIS_all.csv")
write_csv(collapse_data[[3]], "stir_dye_BORIS_all.csv")
write_csv(collapse_data[[4]], "nostir_dye_BORIS_all.csv")

#for the data with aggregated behaviors
write_csv(behav_all[[1]], "chronic_clay_BORIS_agg.csv")
write_csv(behav_all[[2]], "acute_clay_BORIS_agg.csv")
write_csv(behav_all[[3]], "stir_dye_BORIS_agg.csv")
write_csv(behav_all[[4]], "nostir_dye_BORIS_agg.csv")
```