---
title: "Statistical Analysis of Manually Scored Aggressive and Affiliative Behavior"
author: "Hannah Anderson"
date: "2023-04-25"
output: pdf_document
editor_options:
  chunk_output_type: console
---

## Introduction

This file was created to analyze data collected during a series of experiments
conducted from March 30 to August 14, 2022 on the effects of fluctuating
visibility on the behavior of zebrafish (Danio rerio). All data analyzed in this
file was scored using Behavioral Observation Research Interactive Software
(BORIS; Friard & Gamba, 2016) by trained observers. For more information on the
experiments see the file the readme file located in this repository.

The specific behaviors analyzed here are chasing, lateral display and
aggressive contact (aggressive behaviors) and following (affiliative behavior).
For chasing, lateral display and following we scored both the duration of each
behavior and the number of occurrences. Aggressive contact was an instantaneous
behavior and we therefore only scored the number of occurrences.

~will add definitions for each behavior later~

Most of the analyses were conducted using linear mixed modesl with tank ID as a 
random factor to account for multiple recordings of each tank. For the analysis
divided by high/low visibility fluctuations in the acute clay and both dye
experiments we instead used linear models because there were insufficient levels
of tank ID to warrant a random effect.


## Setup

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

setwd("C://Users/hande/Documents/R/fluctuating-turbidity")
library(tidyverse)
library(glmmTMB)
library(emmeans)
library(car)
library(DHARMa)
library(performance)
library(broom)

#importing the base data
chron_clay <- read_csv("chronic_clay_BORIS_all.csv")
clay_dye_all <- read_csv("comparison_clay_dye_BORIS_all_data.csv")
dye_dye_all <- read_csv("comparison_dye_dye_BORIS_all_data.csv")
#also importing the aggregated behavioral data
chron_agg <- read_csv("chronic_clay_BORIS_agg.csv")
clay_dye_agg <- read_csv("comparison_clay_dye_BORIS_agg_data.csv")
dye_dye_agg <- read_csv("comparison_dye_dye_BORIS_agg_data.csv")

#combining the individual and aggregated behavior data into a single dataframe
clay_dye_all <- rbind(clay_dye_all, clay_dye_agg)
dye_dye_all <- rbind(dye_dye_all, dye_dye_agg)

#re-classifying the recording variable as a character for use in the filter()
#function
clay_dye_all$recording <- as.character(clay_dye_all$recording)
dye_dye_all$recording <- as.character(dye_dye_all$recording)

#a list of the scored behaviors to be used in for loops
behavs <- c("chasing", "lat_display", "agg_contact", "following",
            "all_aggressive", "all")


#for the direct comparisons: sub-setting the data by the fourth recording and
#also separating the data by behavior for individual modeling
direct_claydye <- list()
direct_dyedye <- list()
for (i in 1:6) {
  direct_claydye[[i]] <- clay_dye_all %>%
                         filter(recording == "4") %>%
                         filter(behavior == behavs[[i]])
  direct_dyedye[[i]] <- dye_dye_all %>%
                        filter(recording == "4") %>%
                        filter(behavior == behavs[[i]])
}


#for the full comparisons: averaging the two morning recordings together for
#each measure
  #in cases where only one video has a duration value (because the behavior did
  #not occur in the other video) then that value is taken without averaging
morn_data_clay <- clay_dye_all %>%
                  filter(recording %in% c("1", "3")) %>%
                  select(-recording) %>%
                  group_by(shoalID, treatment, experiment, behavior) %>%
                  summarize(duration_s = mean(duration_s, na.rm = TRUE),
                            behav_count = mean(behav_count))
morn_data_dye <- dye_dye_all %>%
                 filter(recording %in% c("1", "3")) %>%
                 select(-recording) %>%
                 group_by(shoalID, treatment, experiment, behavior) %>%
                 summarize(duration_s = mean(duration_s, na.rm = TRUE),
                           behav_count = mean(behav_count))
#labeling these values as the morning values
morn_data_clay$time <- "morning"
morn_data_dye$time <- "morning"
#doing the same with the afternoon values
afte_data_clay <- clay_dye_all %>%
                  filter(recording %in% c("2", "4")) %>%
                  select(-recording) %>%
                  group_by(shoalID, treatment, experiment, behavior) %>%
                  summarize(duration_s = mean(duration_s, na.rm = TRUE),
                            behav_count = mean(behav_count, na.rm = TRUE))
afte_data_dye <- dye_dye_all %>%
                 filter(recording %in% c("2", "4")) %>%
                 select(-recording) %>%
                 group_by(shoalID, treatment, experiment, behavior) %>%
                 summarize(duration_s = mean(duration_s, na.rm = TRUE),
                           behav_count = mean(behav_count, na.rm = TRUE))
afte_data_clay$time <- "afternoon"
afte_data_dye$time <- "afternoon"

#combining the averaged morning and afternoon values into a single dataframe
temp_clay_dye <- rbind(morn_data_clay, afte_data_clay)
temp_dye_dye <- rbind(morn_data_dye, afte_data_dye)

#separating the data by behavior for individual modeling
full_claydye <- list()
full_dyedye <- list()
for (i in 1:6) {
  full_claydye[[i]] <- temp_clay_dye %>%
                       filter(behavior == behavs[[i]])
  full_dyedye[[i]] <- temp_dye_dye %>%
                      filter(behavior == behavs[[i]])
}
```


## Analyses for Experiment 1: Chronic (Two Week) Clay Exposure
#  Analyses divided by behavior and including all recordings

```{r chronic clay all}
#filtering the data by behavior in order to model each behavior separately and
#putting all of them in a list
chronic_list <- list()
for (i in 1:(length(behavs))){
  chronic_list[[i]] <- chron_clay %>%
                     filter(behavior == behavs[i])
}

#models of behavior counts across all recordings
chronic_chase_c <- glmmTMB(behav_count ~ treatment*sex + (1|tankID),
                       data = chronic_list[[1]], family = gaussian())
chronic_lat_c <- update(chronic_chase_c, . ~ ., data = chronic_list[[2]])
chronic_cont_c <- update(chronic_chase_c, . ~ ., data = chronic_list[[3]])
chronic_foll_c <- update(chronic_chase_c, . ~ ., data = chronic_list[[4]])

#checking the model assumptions
plot(simulateResiduals(fittedModel = chronic_chase_c)) #assumptions not met
plot(simulateResiduals(fittedModel = chronic_lat_c)) #assumptions not met
plot(simulateResiduals(fittedModel = chronic_cont_c)) #assumptions not met
plot(simulateResiduals(fittedModel = chronic_foll_c)) #assumptions not met

#transforming the data to meet model assumptions
chron_chase_log <- as.data.frame(chronic_list[[1]])
chron_chase_log$behav_count <- log(chron_chase_log$behav_count+1)
chron_lat_log <- as.data.frame(chronic_list[[2]])
chron_lat_log$behav_count <- log(chron_lat_log$behav_count+1)
chron_cont_log <- as.data.frame(chronic_list[[3]])
chron_cont_log$behav_count <- log(chron_cont_log$behav_count+1)
chron_foll_log <- as.data.frame(chronic_list[[4]])
chron_foll_log$behav_count <- log(chron_foll_log$behav_count+1)

#the tricky transformation (still not working)
test <- update(chronic_lat_c, . ~ ., family = poisson())
plot(simulateResiduals(fittedModel = test))

#the transformed models
chrolog_chase_c <- update(chronic_chase_c, . ~ ., data = chron_chase_log)
chrolog_lat_c <- update(chronic_lat_c, . ~ ., data = chron_lat_log)
chrolog_cont_c <- update(chronic_cont_c, . ~ ., data = chron_cont_log)
chrolog_foll_c <- update(chronic_foll_c, . ~ ., data = chron_foll_log)

#checking the transformed model assumptions
plot(simulateResiduals(fittedModel = chrolog_chase_c))
plot(simulateResiduals(fittedModel = chrolog_lat_c)) #better than untransformed, assumptions still not met
plot(simulateResiduals(fittedModel = chrolog_cont_c))
plot(simulateResiduals(fittedModel = chrolog_foll_c))

#analyzing the results
Anova(chrolog_chase_c)
#Anova(chrolog_lat_c)
Anova(chrolog_cont_c)
Anova(chrolog_foll_c)


#models of behavior durations across all relevant recordings
chronic_chase_d <- update(chronic_chase_c, duration_s ~ .)
chronic_lat_d <- update(chronic_lat_c, duration_s ~ .)
chronic_foll_d <- update(chronic_foll_c, duration_s ~ .)

#checking the model assumptions
plot(simulateResiduals(fittedModel = chronic_chase_d)) #assumptions not met
plot(simulateResiduals(fittedModel = chronic_lat_d)) #assumptions not met
plot(simulateResiduals(fittedModel = chronic_foll_d)) #assumptions not met

#transforming the data to meet model assumptions
chron_chase_log$duration_s <- log(chron_chase_log$duration_s)
chron_lat_log$duration_s <- log(chron_lat_log$duration_s)
chron_foll_log$duration_s <- log(chron_foll_log$duration_s)

#the transformed models
chrolog_chase_d <- update(chronic_chase_d, . ~ ., data = chron_chase_log)
chrolog_lat_d <- update(chronic_lat_d, . ~ ., data = chron_lat_log)
chrolog_foll_d <- update(chronic_foll_d, . ~ ., data = chron_foll_log)

#checking the transformed model assumptions
plot(simulateResiduals(fittedModel = chrolog_chase_d))
plot(simulateResiduals(fittedModel = chrolog_lat_d))
plot(simulateResiduals(fittedModel = chrolog_foll_d))

#analyzing the results
Anova(chrolog_chase_d)
Anova(chrolog_lat_d)
Anova(chrolog_foll_d)
```


## Analyses for Experiment 1: Chronic (Two Week) Clay Exposure
#  Analyses divided by behavior and by high and low visibility fluctuations

```{r chronic clay highlow}
#modelling only the recordings taken when the flux condition had high visibility

#filtering the data by behavior in order to model each behavior separately and
#putting all of them in a list
chronhi_list <- list()
for (i in 1:(length(behavs))){
  chronhi_list[[i]] <- high_chron %>%
                     filter(behavior == behavs[i])
}

#models of behavior counts across high visibility fluctuations
chronhi_chase_c <- glmmTMB(behav_count ~ treatment*sex + (1|tankID),
                       data = chronhi_list[[1]], family = gaussian())
chronhi_lat_c <- update(chronhi_chase_c, . ~ ., data = chronhi_list[[2]])
chronhi_cont_c <- update(chronhi_chase_c, . ~ ., data = chronhi_list[[3]])
chronhi_foll_c <- update(chronhi_chase_c, . ~ ., data = chronhi_list[[4]])

#checking the model assumptions
plot(simulateResiduals(fittedModel = chronhi_chase_c)) #assumptions not met
plot(simulateResiduals(fittedModel = chronhi_lat_c)) #assumptions not met
plot(simulateResiduals(fittedModel = chronhi_cont_c)) #assumptions not met
plot(simulateResiduals(fittedModel = chronhi_foll_c))

#transforming the data to meet model assumptions
chronhi_chase_log <- as.data.frame(chronhi_list[[1]])
chronhi_chase_log$behav_count <- log(chronhi_chase_log$behav_count+1)
chronhi_lat_log <- as.data.frame(chronhi_list[[2]])
chronhi_lat_log$behav_count <- log(chronhi_lat_log$behav_count+1)
chronhi_cont_log <- as.data.frame(chronhi_list[[3]])
chronhi_cont_log$behav_count <- log(chronhi_cont_log$behav_count+1)

#the transformed models
chrologhi_chase_c <- update(chronhi_chase_c, . ~ ., data = chronhi_chase_log)
chrologhi_lat_c <- update(chronhi_lat_c, . ~ ., data = chronhi_lat_log)
chrologhi_cont_c <- update(chronhi_cont_c, . ~ ., data = chronhi_cont_log)

#checking the assumptions of the transformed models
plot(simulateResiduals(fittedModel = chrologhi_chase_c))
plot(simulateResiduals(fittedModel = chrologhi_lat_c))
plot(simulateResiduals(fittedModel = chrologhi_cont_c))

#analyzing the results
Anova(chrologhi_chase_c)
#Anova(chrologhi_lat_c)
Anova(chrologhi_cont_c)
Anova(chronhi_foll_c)


#modelling behavior mean durations across high visibility fluctuations
chronhi_chase_d <- update(chronhi_chase_c, duration_s ~ .)
chronhi_lat_d <- update(chronhi_lat_c, duration_s ~ .)
chronhi_foll_d <- update(chronhi_foll_c, duration_s ~ .)

#checking the assumptions
plot(simulateResiduals(fittedModel = chronhi_chase_d)) #assumptions not met
plot(simulateResiduals(fittedModel = chronhi_lat_d)) #assumptions not met
plot(simulateResiduals(fittedModel = chronhi_foll_d))

#transforming the data to meet model assumptions
chronhi_chase_log$duration_s <- log(chronhi_chase_log$duration_s)
chronhi_lat_log$duration_s <- log(chronhi_lat_log$duration_s)

#the transformed models
chrologhi_chase_d <- update(chronhi_chase_d, . ~ ., data = chronhi_chase_log)
chrologhi_lat_d <- update(chronhi_lat_d, . ~ ., data = chronhi_lat_log)

#checking the transformed model assumptions
plot(simulateResiduals(fittedModel = chrologhi_chase_d))
plot(simulateResiduals(fittedModel = chrologhi_lat_d))

#analyzing the results
Anova(chrologhi_chase_d)
Anova(chrologhi_lat_d)
Anova(chronhi_foll_d)

#post hoc pairwise comparison of the significant result
chrologhi_chase_d %>% emmeans(pairwise ~ treatment)


#modelling only the recordings taken when the flux condition had low visibility

#filtering the data by behavior in order to model each behavior separately and
#putting all of them in a list
chronlo_list <- list()
for (i in 1:(length(behavs))){
  chronlo_list[[i]] <- low_chron %>%
                     filter(behavior == behavs[i])
}

#models of behavior counts across low visibility recordings
chronlo_chase_c <- glmmTMB(behav_count ~ treatment*sex + (1|tankID),
                       data = chronlo_list[[1]], family = gaussian())
chronlo_lat_c <- update(chronlo_chase_c, . ~ ., data = chronlo_list[[2]])
chronlo_cont_c <- update(chronlo_chase_c, . ~ ., data = chronlo_list[[3]])
chronlo_foll_c <- update(chronlo_chase_c, . ~ ., data = chronlo_list[[4]])

#checking the model assumptions
plot(simulateResiduals(fittedModel = chronlo_chase_c)) #assumptions not met
plot(simulateResiduals(fittedModel = chronlo_lat_c)) #assumptions not met
plot(simulateResiduals(fittedModel = chronlo_cont_c))
plot(simulateResiduals(fittedModel = chronlo_foll_c)) #assumptions not met

#transforming the data to meet model assumptions
chronlo_chase_log <- as.data.frame(chronlo_list[[1]])
chronlo_chase_log$behav_count <- log(chronlo_chase_log$behav_count+1)
chronlo_lat_log <- as.data.frame(chronlo_list[[2]])
chronlo_lat_log$behav_count <- log(chronlo_lat_log$behav_count+1)
chronlo_foll_log <- as.data.frame(chronlo_list[[4]])
chronlo_foll_log$behav_count <- log(chronlo_foll_log$behav_count+1)

#the transformed models
chrologlo_chase_c <- update(chronlo_chase_c, . ~ ., data = chronlo_chase_log)
chrologlo_lat_c <- update(chronlo_lat_c, . ~ ., data = chronlo_lat_log)
chrologlo_foll_c <- update(chronlo_foll_c, . ~ ., data = chronlo_foll_log)

#checking the transformed model assumptions
plot(simulateResiduals(fittedModel = chrologlo_chase_c))
plot(simulateResiduals(fittedModel = chrologlo_lat_c)) #assumptions not met
plot(simulateResiduals(fittedModel = chrologlo_foll_c))

#analyzing the results
Anova(chrologlo_chase_c)
#Anova(chrologlo_lat_c)
Anova(chronlo_cont_c)
Anova(chrologlo_foll_c)

#post hoc pairwise comparison of the significant result
chrologlo_foll_c %>% emmeans(pairwise ~ treatment)


#modelling  behavior mean durations across low visibility fluctuations
chronlo_chase_d <- update(chronlo_chase_c, duration_s ~ .)
chronlo_lat_d <- update(chronlo_lat_c, duration_s ~ .)
chronlo_foll_d <- update(chronlo_foll_c, duration_s ~ .)

#checking the model assumptions
plot(simulateResiduals(fittedModel = chronlo_chase_d)) #assumptions not met
plot(simulateResiduals(fittedModel = chronlo_lat_d)) #assumptions not met
plot(simulateResiduals(fittedModel = chronlo_foll_d)) #assumptions not met

#transforming the data to meet model assumptions
chronlo_chase_log <- as.data.frame(chronlo_list[[1]])
chronlo_chase_log$duration_s <- log(chronlo_chase_log$duration_s)
chronlo_lat_log$duration_s <- log(chronlo_lat_log$duration_s)
chronlo_foll_log <- as.data.frame(chronlo_list[[4]])
chronlo_foll_log$duration_s <- log(chronlo_foll_log$duration_s)

#the transformed models
chrologlo_chase_d <- update(chronlo_chase_d, . ~ ., data = chronlo_chase_log)
chrologlo_lat_d <- update(chronlo_lat_d, . ~ ., data = chronlo_lat_log)
chrologlo_foll_d <- update(chronlo_foll_d, . ~ ., data = chronlo_foll_log)

#checking the assumptions of the transformed models
plot(simulateResiduals(fittedModel = chrologlo_chase_d))
plot(simulateResiduals(fittedModel = chrologlo_lat_d))
plot(simulateResiduals(fittedModel = chrologlo_foll_d))

#analyzing the results
Anova(chrologlo_chase_d)
Anova(chrologlo_lat_d)
Anova(chrologlo_foll_d)
```


## Direct Comparisons of Count Data

```{r direct count}
#the clay vs dye models

#chases
d_con_cd_chase <- lm(behav_count ~ treatment*experiment,
                     data = direct_claydye[[1]])
#lateral displays
d_con_cd_lat <- update(d_con_cd_chase, . ~ ., data = direct_claydye[[2]])
#aggressive contacts
d_con_cd_cont <- update(d_con_cd_chase, . ~ ., data = direct_claydye[[3]])
#follows
d_con_cd_foll <- update(d_con_cd_cont, . ~ ., data = direct_claydye[[4]])
#aggregated aggressive behaviors
d_con_cd_angr <- update(d_con_cd_chase, . ~ ., data = direct_claydye[[5]])
#all behaviors aggregated
d_con_cd_all <- update(d_con_cd_chase, . ~ ., data = direct_claydye[[6]])

#checking the model assumptions
#plot(d_con_cd_chase) #assumptions not met
#plot(d_con_cd_lat) #assumptions not met
#plot(d_con_cd_cont)
#plot(d_con_cd_foll)
#plot(d_con_cd_angr)
#plot(d_con_cd_all) #assumptions borderline

#transforming the data to meet model assumptions
direct_cd_trans <- direct_claydye
direct_cd_trans[[1]]$behav_count <- sqrt(direct_cd_trans[[1]]$behav_count) +
                                      sqrt(direct_cd_trans[[1]]$behav_count+1)
direct_cd_trans[[2]]$behav_count <- sqrt(direct_cd_trans[[2]]$behav_count) +
                                      sqrt(direct_cd_trans[[2]]$behav_count+1)
direct_cd_trans[[6]]$behav_count <- sqrt(direct_cd_trans[[6]]$behav_count) +
                                      sqrt(direct_cd_trans[[6]]$behav_count+1)

#the transformed models
d_con_cd_chase_trans <- update(d_con_cd_chase, . ~ .,
                               data = direct_cd_trans[[1]])
d_con_cd_lat_trans <- update(d_con_cd_lat, . ~ ., data = direct_cd_trans[[2]])
d_con_cd_all_trans <- update(d_con_cd_all, . ~ ., data = direct_cd_trans[[6]])

#checking the transformed model assumptions
#plot(d_con_cd_chase_trans)
#plot(d_con_cd_lat_trans)
#plot(d_con_cd_all_trans)


#the dye vs dye models

#chases
d_con_dd_chase <- lm(behav_count ~ treatment*experiment,
                     data = direct_dyedye[[1]])
#lateral displays
d_con_dd_lat <- update(d_con_dd_chase, . ~ ., data = direct_dyedye[[2]])
#aggressive contacts
d_con_dd_cont <- update(d_con_dd_chase, . ~ ., data = direct_dyedye[[3]])
#follows
d_con_dd_foll <- update(d_con_dd_chase, . ~ ., data = direct_dyedye[[4]])
#aggregated aggressive behaviors
d_con_dd_angr <- update(d_con_dd_chase, . ~ ., data = direct_dyedye[[5]])
#all behaviors aggregated
d_con_dd_all <- update(d_con_dd_chase, . ~ ., data = direct_dyedye[[6]])

#checking the model assumptions
#plot(d_con_dd_chase)
#plot(d_con_dd_lat) #assumptions not met
#plot(d_con_dd_cont) #assumptions not met
#plot(d_con_dd_foll)
#plot(d_con_dd_angr)
#plot(d_con_dd_all) #assumptions not met

#transforming the data to meet model assumptions
direct_dd_trans <- direct_dyedye
direct_dd_trans[[2]]$behav_count <- sqrt(direct_dd_trans[[2]]$behav_count) +
                                    sqrt(direct_dd_trans[[2]]$behav_count+1)
direct_dd_trans[[3]]$behav_count <- sqrt(direct_dd_trans[[3]]$behav_count) +
                                    sqrt(direct_dd_trans[[3]]$behav_count+1)
direct_dd_trans[[6]]$behav_count <- sqrt(direct_dd_trans[[6]]$behav_count) +
                                    sqrt(direct_dd_trans[[6]]$behav_count+1)

#the transformed models
d_con_dd_lat_trans <- update(d_con_dd_lat, . ~ ., data = direct_dd_trans[[2]])
d_con_dd_cont_trans <- update(d_con_dd_cont, . ~ ., data = direct_dd_trans[[3]])
d_con_dd_all_trans <- update(d_con_dd_all, . ~ ., data = direct_dd_trans[[6]])

#checking the transformed model assumptions
#plot(d_con_dd_lat_trans)
#plot(d_con_dd_cont_trans)
#plot(d_con_dd_all_trans)


#analyzing the results of the LMs and putting in tibbles within a list for easy
#export
results_lm <- list()
results_lm[[1]] <- tidy(Anova(d_con_cd_chase_trans))
results_lm[[2]] <- tidy(Anova(d_con_cd_lat_trans))
results_lm[[3]] <- tidy(Anova(d_con_cd_cont))
results_lm[[4]] <- tidy(Anova(d_con_cd_foll))
results_lm[[5]] <- tidy(Anova(d_con_cd_angr))
results_lm[[6]] <- tidy(Anova(d_con_cd_all_trans))
results_lm[[7]] <- tidy(Anova(d_con_dd_chase))
results_lm[[8]] <- tidy(Anova(d_con_dd_lat_trans))
results_lm[[9]] <- tidy(Anova(d_con_dd_cont_trans))
results_lm[[10]] <- tidy(Anova(d_con_dd_foll))
results_lm[[11]] <- tidy(Anova(d_con_dd_angr))
results_lm[[12]] <- tidy(Anova(d_con_dd_all_trans))

#adding columns with additional identifying information to all tibbles within
#the results list
dependent_variables <- c("chasing", "lateral_display", "aggressive_contact",
                         "following", "aggregated_aggression", "all_behaviors",
                         "chasing", "lateral_display", "aggressive_contact",
                         "following", "aggregated_aggression", "all_behaviors")
for (i in 1:6) {
  results_lm[[i]]$experiments <- "claydye"
  results_lm[[i]]$dependent_variable <- dependent_variables[[i]]
}
for (i in 7:12) {
  results_lm[[i]]$experiments <- "dyedye"
  results_lm[[i]]$dependent_variable <- dependent_variables[[i]]
}
#combining the LM results list into a single tibble
results_lm_coll <- do.call(rbind, results_lm)
#reordering the columns to be more intuitive
results_lm_coll <- results_lm_coll %>%
                   select(experiments, dependent_variable, term, sumsq, df,
                          statistic, p.value)

#writing the results to .csv
write_csv(results_lm_coll, "direct_comparison_BORIS_count_LMs.csv")
```


## Direct Comparison of Duration Data

```{r direct duration}
#the clay vs dye models

#chases
d_dur_cd_chase <- lm(duration_s ~ treatment*experiment,
                     data = direct_claydye[[1]])
#lateral displays
d_dur_cd_lat <- update(d_dur_cd_chase, . ~ ., data = direct_claydye[[2]])
#follows
d_dur_cd_foll <- update(d_dur_cd_chase, . ~ ., data = direct_claydye[[4]])
#aggregated aggressive behaviors
d_dur_cd_angr <- update(d_dur_cd_chase, . ~ ., data = direct_claydye[[5]])
#all behaviors aggregated
d_dur_cd_all <- update(d_dur_cd_chase, . ~ ., data = direct_claydye[[6]])

#checking the model assumptions
#plot(d_dur_cd_chase) #assumptions not met
#plot(d_dur_cd_lat)
#plot(d_dur_cd_foll)
#plot(d_dur_cd_angr) #assumptions not met
#plot(d_dur_cd_all)

#transforming the data to meet model assumptions
direct_cd_trans <- direct_claydye
direct_cd_trans[[1]]$duration_s <- log(direct_cd_trans[[1]]$duration_s)
direct_cd_trans[[5]]$duration_s <- log(direct_cd_trans[[5]]$duration_s)

#the transformed models
d_dur_cd_chase_trans <- update(d_dur_cd_chase, . ~ .,
                               data = direct_cd_trans[[1]])
d_dur_cd_angr_trans <- update(d_dur_cd_angr, . ~ ., data = direct_cd_trans[[5]])

#checking the transformed model assumptions
#plot(d_dur_cd_chase_trans)
#plot(d_dur_cd_angr_trans)


#the dye vs dye models

#chases
d_dur_dd_chase <- lm(duration_s ~ treatment*experiment,
                     data = direct_dyedye[[1]])
#lateral displays
d_dur_dd_lat <- update(d_dur_dd_chase, . ~ ., data = direct_dyedye[[2]])
#follows
d_dur_dd_foll <- update(d_dur_dd_chase, . ~ ., data = direct_dyedye[[4]])
#aggregated aggressive behaviors
d_dur_dd_angr <- update(d_dur_dd_chase, . ~ ., data = direct_dyedye[[5]])
#all behaviors aggregated
d_dur_dd_all <- update(d_dur_dd_chase, . ~ ., data = direct_dyedye[[6]])

#checking the model assumptions
#plot(d_dur_dd_chase)
#plot(d_dur_dd_lat)
#plot(d_dur_dd_foll) #assumptions not met
#plot(d_dur_dd_angr)
#plot(d_dur_dd_all) #assumptions not met

#transforming the data to meet model assumptions
direct_dd_trans <- direct_dyedye
direct_dd_trans[[4]]$duration_s <- log(direct_dd_trans[[4]]$duration_s)
direct_dd_trans[[6]]$duration_s <- log(direct_dd_trans[[6]]$duration_s)

#the transformed model
d_dur_dd_chase_trans <- update(d_dur_dd_chase, . ~ .,
                               data = direct_dd_trans[[4]])
d_dur_dd_all_trans <- update(d_dur_dd_all, . ~ ., data = direct_dd_trans[[6]])

#checking the transformed model assumptions
#plot(d_dur_dd_chase_trans)
#plot(d_dur_dd_all_trans)


#analyzing the results of the LMs and putting them in tibbles within a list for
#easy export
results_lm <- list()
results_lm[[1]] <- tidy(Anova(d_dur_cd_chase_trans))
results_lm[[2]] <- tidy(Anova(d_dur_cd_lat))
results_lm[[3]] <- tidy(Anova(d_dur_cd_foll))
results_lm[[4]] <- tidy(Anova(d_dur_cd_angr_trans))
results_lm[[5]] <- tidy(Anova(d_dur_cd_all))
results_lm[[6]] <- tidy(Anova(d_dur_dd_chase_trans))
results_lm[[7]] <- tidy(Anova(d_dur_dd_lat))
results_lm[[8]] <- tidy(Anova(d_dur_dd_foll))
results_lm[[9]] <- tidy(Anova(d_dur_dd_angr))
results_lm[[10]] <- tidy(Anova(d_dur_dd_all_trans))

#adding columns with additional identifying information to all tibbles within
#the results list
dependent_variables <- c("chasing", "lateral_display", "following",
                         "aggregated_aggression", "all_behaviors",
                         "chasing", "lateral_display", "following",
                         "aggregated_aggression", "all_behaviors")
for (i in (1:5)) {
  results_lm[[i]]$experiments <- "claydye"
  results_lm[[i]]$dependent_variable <- dependent_variables[[i]]
}
for (i in 6:10) {
  results_lm[[i]]$experiments <- "dyedye"
  results_lm[[i]]$dependent_variable <- dependent_variables[[i]]
}
#combining the LM results list into a single tibble
results_lm_coll <- do.call(rbind, results_lm)
#reordering the columns to be more intuitive
results_lm_coll <- results_lm_coll %>%
                   select(experiments, dependent_variable, term, sumsq, df,
                          statistic, p.value)

#exporting the results to .csv
write_csv(results_lm_coll, "direct_comparison_BORIS_duration_LMs.csv")
```


## Full Comparisons of Count Data

```{r full count}
#the clay vs dye models

#chases
f_con_cd_chase <- glmmTMB(behav_count ~ treatment*experiment*time + (1|shoalID),
                          data = full_claydye[[1]], family = gaussian())
#lateral displays
f_con_cd_lat <- update(f_con_cd_chase, . ~ ., data = full_claydye[[2]])
#aggressive contacts
f_con_cd_cont <- update(f_con_cd_chase, . ~ ., data = full_claydye[[3]])
#follows
f_con_cd_foll <- update(f_con_cd_chase, . ~ ., data = full_claydye[[4]])
#aggregated aggressive behaviors
f_con_cd_angr <- update(f_con_cd_chase, . ~ ., data = full_claydye[[5]])
#all behaviors aggregated
f_con_cd_all <- update(f_con_cd_chase, . ~ ., data = full_claydye[[6]])

#checking the model assumptions
#plot(simulateResiduals(fittedModel = f_con_cd_chase))
#plot(simulateResiduals(fittedModel = f_con_cd_lat)) #assumptions not met
#plot(simulateResiduals(fittedModel = f_con_cd_cont))
#plot(simulateResiduals(fittedModel = f_con_cd_foll)) #assumptions not met
#plot(simulateResiduals(fittedModel = f_con_cd_angr))
#plot(simulateResiduals(fittedModel = f_con_cd_all))

#transforming the data to meet model assumptions
full_cd_trans <- full_claydye
full_cd_trans[[2]]$behav_count <- sqrt(full_cd_trans[[2]]$behav_count) +
                                  sqrt(full_cd_trans[[2]]$behav_count+1)
full_cd_trans[[4]]$behav_count <- sqrt(full_cd_trans[[4]]$behav_count) +
                                  sqrt(full_cd_trans[[4]]$behav_count+1)

#the transformed models
f_con_cd_lat_trans <- update(f_con_cd_lat, . ~ ., data = full_cd_trans[[2]])
f_con_cd_foll_trans <- update(f_con_cd_foll, . ~ ., data = full_cd_trans[[4]])

#testing the model assumptions
#plot(simulateResiduals(fittedModel = f_con_cd_lat_trans))
#plot(simulateResiduals(fittedModel = f_con_cd_foll_trans))




#the dye vs dye models

#chases
f_con_dd_chase <- glmmTMB(behav_count ~ treatment*experiment*time + (1|shoalID),
                          data = full_dyedye[[1]], family = gaussian())
#lateral displays
f_con_dd_lat <- update(f_con_cd_chase, . ~ ., data = full_dyedye[[2]])
#aggressive contacts
f_con_dd_cont <- update(f_con_cd_chase, . ~ ., data = full_dyedye[[3]])
#follows
f_con_dd_foll <- update(f_con_cd_chase, . ~ ., data = full_dyedye[[4]])
#aggregated aggressive behaviors
f_con_dd_angr <- update(f_con_cd_chase, . ~ ., data = full_dyedye[[5]])
#all behaviors aggregated
f_con_dd_all <- update(f_con_cd_chase, . ~ ., data = full_dyedye[[6]])

#checking the model assumptions
#plot(simulateResiduals(fittedModel = f_con_dd_chase))
#plot(simulateResiduals(fittedModel = f_con_dd_lat)) #assumptions not met
#plot(simulateResiduals(fittedModel = f_con_dd_cont))
#plot(simulateResiduals(fittedModel = f_con_dd_foll)) #outlier
#plot(simulateResiduals(fittedModel = f_con_dd_angr))
#plot(simulateResiduals(fittedModel = f_con_dd_all))

#transforming the data to meet model assumptions
full_dd_trans <- full_dyedye
full_dd_trans[[2]]$behav_count <- sqrt(full_dd_trans[[2]]$behav_count) +
                                  sqrt(full_dd_trans[[2]]$behav_count+1)
full_dd_trans[[4]]$behav_count <- sqrt(full_dd_trans[[4]]$behav_count) +
                                  sqrt(full_dd_trans[[4]]$behav_count+1)

#the transformed model
f_con_dd_lat_trans <- update(f_con_dd_lat, . ~ ., data = full_dd_trans[[2]])
f_con_dd_foll_trans <- update(f_con_dd_foll, . ~ ., data = full_dd_trans[[4]])

#checking the model assumptions
#plot(simulateResiduals(fittedModel = f_con_dd_lat_trans))
#plot(simulateResiduals(fittedModel = f_con_dd_foll_trans))


#analyzing the results of the LMMs and putting them in tibbles within a list for
#easy export
results_lmm <- list()
results_lmm[[1]] <- tidy(Anova(f_con_cd_chase))
results_lmm[[2]] <- tidy(Anova(f_con_cd_lat_trans))
results_lmm[[3]] <- tidy(Anova(f_con_cd_cont))
results_lmm[[4]] <- tidy(Anova(f_con_cd_foll_trans))
results_lmm[[5]] <- tidy(Anova(f_con_cd_angr))
results_lmm[[6]] <- tidy(Anova(f_con_cd_all))
results_lmm[[7]] <- tidy(Anova(f_con_dd_chase))
results_lmm[[8]] <- tidy(Anova(f_con_dd_lat_trans))
results_lmm[[9]] <- tidy(Anova(f_con_dd_cont))
results_lmm[[10]] <- tidy(Anova(f_con_dd_foll_trans))
results_lmm[[11]] <- tidy(Anova(f_con_dd_angr))
results_lmm[[12]] <- tidy(Anova(f_con_dd_all))

#adding columns with additional identifying information to all tibbles within
#the results list
dependent_variables <- c("chasing", "lateral_display", "aggressive_contact",
                         "following", "aggregated_aggression", "all_behaviors",
                         "chasing", "lateral_display", "aggressive_contact",
                         "following", "aggregated_aggression", "all_behaviors")
for (i in 1:6) {
  results_lmm[[i]]$experiments <- "claydye"
  results_lmm[[i]]$dependent_variable <- dependent_variables[[i]]
}
for (i in 7:12) {
  results_lmm[[i]]$experiments <- "dyedye"
  results_lmm[[i]]$dependent_variable <- dependent_variables[[i]]
}
#combining the LMM results list into a single tibble
results_lmm_coll <- do.call(rbind, results_lmm)
#reordering the columns to be more intuitive
results_lmm_coll <- results_lmm_coll %>%
                    select(experiments, dependent_variable, term, df, statistic,
                           p.value)

#post hoc pairwise comparisons of the significant results
pair_lmm <- list()
pair_lmm[[1]] <- as.data.frame(pairs(emmeans(f_con_cd_cont, ~ experiment |
                                                              time)))
pair_lmm[[2]] <- as.data.frame(pairs(emmeans(f_con_cd_angr, ~ experiment |
                                                              time)))
pair_lmm[[3]] <- as.data.frame(pairs(emmeans(f_con_cd_all, ~ experiment |
                                                             time)))

for (i in 1:3) {
  pair_lmm[[i]]$experiments <- "claydye"
}
pair_lmm[[1]]$dependent_variable <- "aggressive_contact"
pair_lmm[[2]]$dependent_variable <- "aggregated_aggression"
pair_lmm[[3]]$dependent_variable <- "all_behaviors"
pair_lmm_coll <- do.call(rbind, pair_lmm)
pair_lmm_coll <- pair_lmm_coll %>%
                 select(experiments, dependent_variable, everything())

#exporting the results to .csv
write_csv(results_lmm_coll, "full_comparison_BORIS_count_LMMs.csv")
write_csv(pair_lmm_coll, "full_pairwise_comparison_BORIS_count_LMMs.csv")
```


## Full Comparisons of Duration Data

```{r full duration}
#the clay vs dye models

#chases
f_dur_cd_chase <- glmmTMB(duration_s ~ treatment*experiment*time + (1|shoalID),
                          data = full_claydye[[1]], family = gaussian())
#lateral displays
f_dur_cd_lat <- update(f_dur_cd_chase, . ~ ., data = full_claydye[[2]])
#follows
f_dur_cd_foll <- update(f_dur_cd_chase, . ~ ., data = full_claydye[[4]])
#aggregated aggressive behaviors
f_dur_cd_angr <- update(f_dur_cd_chase, . ~ ., data = full_claydye[[5]])
#all behaviors aggregated
f_dur_cd_all <- update(f_dur_cd_chase, . ~ ., data = full_claydye[[6]])

#checking the model assumptions
#plot(simulateResiduals(fittedModel = f_dur_cd_chase))
#plot(simulateResiduals(fittedModel = f_dur_cd_lat)) #outliers
#plot(simulateResiduals(fittedModel = f_dur_cd_foll))
#plot(simulateResiduals(fittedModel = f_dur_cd_angr))
#plot(simulateResiduals(fittedModel = f_dur_cd_all))


#the dye/dye comparison models

#chases
f_dur_dd_chase <- glmmTMB(duration_s ~ treatment*experiment*time + (1|shoalID),
                          data = full_dyedye[[1]], family = gaussian())
#lateral displays
f_dur_dd_lat <- update(f_dur_dd_chase, . ~ ., data = full_dyedye[[2]])
#follows
f_dur_dd_foll <- update(f_dur_dd_chase, . ~ ., data = full_dyedye[[4]])
#aggregated aggressive behaviors
f_dur_dd_angr <- update(f_dur_dd_chase, . ~ ., data = full_dyedye[[5]])
#all behaviors aggregated
f_dur_dd_all <- update(f_dur_dd_chase, . ~ ., data = full_dyedye[[6]])

#checking the model assumptions
#plot(simulateResiduals(fittedModel = f_dur_dd_chase)) #assumptions not met
#plot(simulateResiduals(fittedModel = f_dur_dd_lat))
#plot(simulateResiduals(fittedModel = f_dur_dd_foll))
#plot(simulateResiduals(fittedModel = f_dur_dd_angr)) #assumptions not met
#plot(simulateResiduals(fittedModel = f_dur_dd_all))

#transforming the data to meet model assumptions
full_dd_trans <- full_dyedye
full_dd_trans[[1]]$duration_s <- 1/(full_dd_trans[[1]]$duration_s)
full_dd_trans[[5]]$duration_s <- log(full_dd_trans[[5]]$duration_s)

#the transformed models
f_dur_dd_chase_trans <- update(f_dur_dd_chase, . ~ ., data = full_dd_trans[[1]])
f_dur_dd_angr_trans <- update(f_dur_dd_angr, . ~ ., data = full_dd_trans[[5]])

#checking the transformed model assumptions
#plot(simulateResiduals(fittedModel = f_dur_dd_chase_trans))
#plot(simulateResiduals(fittedModel = f_dur_dd_angr_trans))


#analyzing the results of the LMMs and putting them in tibbles within a list for
#easy export
results_lmm <- list()
results_lmm[[1]] <- tidy(Anova(f_dur_cd_chase))
results_lmm[[2]] <- tidy(Anova(f_dur_cd_lat))
results_lmm[[3]] <- tidy(Anova(f_dur_cd_foll))
results_lmm[[4]] <- tidy(Anova(f_dur_cd_angr))
results_lmm[[5]] <- tidy(Anova(f_dur_cd_all))
results_lmm[[6]] <- tidy(Anova(f_dur_dd_chase_trans))
results_lmm[[7]] <- tidy(Anova(f_dur_dd_lat))
results_lmm[[8]] <- tidy(Anova(f_dur_dd_foll))
results_lmm[[9]] <- tidy(Anova(f_dur_dd_angr_trans))
results_lmm[[10]] <- tidy(Anova(f_dur_dd_all))

#adding columns with additional identifying information to all tibbles within
#the results list
dependent_variables <- c("chasing", "lateral_display", "following",
                         "aggregated_aggression", "all_behaviors",
                         "chasing", "lateral_display", "following",
                         "aggregated_aggression", "all_behaviors")
for (i in 1:5) {
  results_lmm[[i]]$experiments <- "claydye"
  results_lmm[[i]]$dependent_variable <- dependent_variables[[i]]
}
for (i in 6:10) {
  results_lmm[[i]]$experiments <- "dyedye"
  results_lmm[[i]]$dependent_variable <- dependent_variables[[i]]
}
#combining the LMM results list into a single tibble
results_lmm_coll <- do.call(rbind, results_lmm)
results_lmm_coll <- results_lmm_coll %>%
                    select(experiments, dependent_variable, term, df, statistic,
                           p.value)

#post hoc pairwise comparisons of the significant LMM results
pair_lmm <- list()
pair_lmm[[1]] <- as.data.frame(pairs(emmeans(f_dur_cd_foll, ~ treatment)))
pair_lmm[[2]] <- as.data.frame(pairs(emmeans(f_dur_dd_angr_trans, ~ treatment |
                                                                    time)))
pair_lmm[[3]] <- as.data.frame(pairs(emmeans(f_dur_dd_all, ~ treatment)))

pair_lmm[[1]]$experiments <- "claydye"
for (i in 2:3) {
  pair_lmm[[i]]$experiments <- "dyedye"
}
pair_lmm[[1]]$dependent_variable <- "following"
pair_lmm[[2]]$dependent_variable <- "aggregated_aggression"
pair_lmm[[3]]$dependent_variable <- "all_behaviors"
for (i in 1:3) {
  pair_lmm[[i]] <- pair_lmm[[i]] %>%
                   select(experiments, dependent_variable, everything())
}
pair_lmm_coll <- rbind(pair_lmm[[1]], pair_lmm[[3]])

#exporting the results to .csv
write_csv(results_lmm_coll, "full_comparison_BORIS_duration_LMMs.csv")
write_csv(pair_lmm_coll, "full_pairwise_oneway_comparisons_duration_LMMs.csv")
write_csv(pair_lmm[[2]], "full_pairwise_all_behaviors_duration_LMMs.csv")
```