---
title: "Chronic Clay Statistics"
author: "Hannah Anderson"
date: "2024-06-05"
output: pdf_document
editor_options: 
  chunk_output_type: console
---

## Introduction

These are the statistical analyses for the manuscript [manuscript ID here].
These analyses include three(?) experiments. In brief, these experiments were
conducted to observe the behavior of shoals of zebrafish under control
visibility (clear), consistently poor visibility (stable) and fluctuating
visibility (flux). All three shared the same basic treatment conditions, though
behavioral measures and the rate of fluctuation (in the flux condition) varied.

Additional info here.

See readme for full explanation of methods.



```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

setwd("C://Users/hande/Documents/R/fluctuating-turbidity")
library(tidyverse)
library(glmmTMB)
library(emmeans)
library(car)
library(DHARMa)
library(broom)

#importing the data
all_chronic_proxim <- read_csv("chronic_clay_social_data.csv")
all_chronic_move <- read_csv("chronic_clay_movement_data.csv")
#all_chronic_boris <- read_csv("chronic_clay_BORIS_all.csv")
```


## Proximity Analyses

-full comparison stats of entire chronic clay experiment
-direct comparison stats of entire chronic clay experiment
  -for both of the above I won't be using the afternoon recordings from
   day 1 and 2
-acute analysis of behavioral changes over first two days/fluctuations
  -will use all four recordings from day 1 and 2

```{r proximity}
###Chronic Analyses

#filtering out the afternoon recordings from day 1 and 2 for consistent
#recording time (morning)
chronic_proxim <- all_chronic_proxim %>%
                  filter(recording != "rec2")
chronic_proxim <- chronic_proxim %>%
                  filter(recording != "rec4")
#re-numbering the recordings for the chronic/long-term analysis to correspond to
#the day of the recording rather than the order
chronic_proxim <- chronic_proxim %>%
                  mutate(recording = case_match(recording,
                                                "rec1" ~ 1,
                                                "rec3" ~ 2,
                                                "rec5" ~ 7,
                                                "rec7" ~ 8,
                                                "rec8" ~ 13,
                                                "rec9" ~ 14))
chronic_proxim <- chronic_proxim %>%
                  rename_at("recording", ~"recording_day")


##Full Comparison Models

#mean shoal area
f_shoal <- lm(shoal_area ~ treatment*recording_day*sex, data = chronic_proxim)
#mean neighbor distance
f_neigh <- update(f_shoal, mean_neighbor ~ .)
#polarity
f_polar <- update(f_shoal, polarity ~ .)

#checking the model assumptions
plot(simulateResiduals(fittedModel = f_shoal)) #assumptions not met
plot(simulateResiduals(fittedModel = f_neigh))
plot(simulateResiduals(fittedModel = f_polar)) #assumptions not met

#transforming the shoal area data to meet model assumptions
trans_proxim <- chronic_proxim
trans_proxim$shoal_area <- log(chronic_proxim$shoal_area)
trans_proxim$polarity <- asin(sqrt(trans_proxim$polarity))

#the transformed model
f_trans_shoal <- update(f_shoal, . ~ ., data = trans_proxim)

#checking the transformed model assumptions
plot(simulateResiduals(fittedModel = f_trans_shoal))

#using an alternate family model to meet polarity model assumptions
f_beta_polar <- glmmTMB(polarity ~ treatment*recording_day*sex,
                        data = chronic_proxim,
                        family = beta_family(link = "logit"))

#checking the GLM model assumptions
plot(simulateResiduals(fittedModel = f_beta_polar))


#analyzing the results of the LMs and putting them in tibbles within a list for
#easy export
results_lm <- list()
results_lm[[1]] <- tidy(Anova(f_trans_shoal)) #shoal area
results_lm[[2]] <- tidy(Anova(f_neigh)) #neighbor distance

#adding a column to identify the measure
results_lm[[1]]$dependent_variable <- "shoal_area"
results_lm[[2]]$dependent_variable <- "neighbor_distance"

#combining the LM results list into a single tibble
results_lm_coll <- rbind(results_lm[[1]], results_lm[[2]])

#analyzing the results of the GLM and placing in a tibble
results_glm <- tidy(Anova(f_beta_polar))
f_beta_polar$dependent_variable <- "polarity"

#exporting the results to .csv
write_csv(results_lm_coll, "full_chronic_proximity_LMs.csv")
write_csv(results_glm, "full_chronic_proximity_GLM.csv")



##Direct Comparison Models

#sub-setting the data to only include recordings where the stable and flux
#conditions are the same visibility level for the direct comparisons
direct_proxim <- chronic_proxim %>%
                 filter(recording_day == 2 |
                        recording_day == 8 |
                        recording_day == 14)

#mean shoal area
d_shoal <- lm(shoal_area ~ treatment*recording_day*sex, data = direct_proxim)
#mean neighbor distance
d_neigh <- update(d_shoal, mean_neighbor ~ .)
#polarity
d_polar <- update(d_shoal, polarity ~ .)

plot(simulateResiduals(fittedModel = d_shoal))
plot(simulateResiduals(fittedModel = d_neigh))
plot(simulateResiduals(fittedModel = d_polar))



###Acute Analyses

#sub-setting the data to only include the first four recordings to observe
#the acute effects of treatment
acute_proxim <- all_chronic_proxim %>%
                filter(recording == c("rec1", "rec2", "rec3", "rec4"))
#converting the recording ID from a string to a numeric for analysis purposes
acute_proxim <- acute_proxim %>%
                mutate(recording = case_match(recording,
                                              "rec1" ~ 1,
                                              "rec2" ~ 2,
                                              "rec3" ~ 3,
                                              "rec4" ~ 4))

#mean shoal area
a_shoal <- lm(shoal_area ~ treatment*recording*sex, data = acute_proxim)
#mean neighbor distance
a_neigh <- update(d_shoal, mean_neighbor ~ .)
#polarity
a_polar <- update(d_shoal, polarity ~ .)

#checking the model assumptions
plot(simulateResiduals(fittedModel = a_shoal)) #assumptions not met
plot(simulateResiduals(fittedModel = a_neigh))
plot(simulateResiduals(fittedModel = a_polar))

#transforming the data to meet model assumptions
trans_proxim <- acute_proxim
trans_proxim$shoal_area <- log(trans_proxim$shoal_area)

#the transformed model
a_trans_shoal <- update(a_shoal, . ~ ., data = trans_proxim)

#checking the transformed model assumptions
plot(simulateResiduals(fittedModel = a_trans_shoal))
```
