---
title: "Fluctuating Turbidity Data Conversion for Sociality Tests"
author: "Hannah Anderson"
date: "2023-02-03"
output: pdf_document
---

## Introduction

Data for this experiment was captured using vertically mounted cameras
observing the movement of zebrafish in shallow waters. The videos were then
processed using the automated tracking program TRex, which returns individual
files for each tracked fish for each video. Here, we are taking these individual
fish files and 1) converting the frame-by-frame data into summary statistics
for each video and 2) combining the data from the 500+ files into a singular
file. This file is used to export social data. See [insert filename] for the
conversion of movement data. If using one file after the other make sure to
clear the working environment as many variables are shared across files.

This file will be used to export movement data. See [insert filename] for
the conversion of social and shoaling data. If using one after the other make
sure to clear the working environment as many variables are shared across files.


## Setup

For our experiments fish occurred in either 3- or 4-fish groups. Change the
value of "fish" in the following chunk to account for the number of fish in the
experiment currently being analyzed.

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

#setting the working directory
knitr::opts_knit$set(root.dir = "D://2 Day Mini Kaolin Turb Flux Experiment/testing")
library(tidyverse)
library(data.table)
library(zoo)
#the number of fish per group
fish = 4
```

Setting up the code to be able to read in all of the files from a given
directory. We are also changing the "Inf" values TRex uses in place of NAs to
NAs and creating a master data frame to contain all of the summary statistics.

```{r file read in, message = FALSE}
#reading the file names into a list
list_csv <- list.files(pattern = "*.csv")
#from the list, creating a list of dataframes while excluding unused columns
masterlist <- lapply(list_csv, read_csv, col_select = c(3:4, 6:8))
#replacing the inf TRex uses in place of NAs with NAs
Inf2NA <- function(v) replace(v, is.infinite(v), NA)
fishlist <- lapply(masterlist, function(d) replace(d, TRUE, sapply(d, Inf2NA)))
```

## Processing Files

The below code uses nested for-loops to process each of the datafiles and
eventually combine them into a single master file. For each video, TRex creates
a datafile for each detected fish- in the case of these experiments, that means
three or four fish per video (depending on the experiment). For analysis we are
lumping all fish behaviors across a group because individual identification is
impossible.

In this file we are extracting social data. We do this by... something about
X,Y coordinates and angles.

```{r loop, include = FALSE, message = FALSE}
plot(pressure)
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
