---
title: "Fluctuating Turbidity Data Conversion for Sociality Tests"
author: "Hannah Anderson"
date: "2023-02-03"
output: pdf_document
---

## Introduction

Data for this experiment was captured using vertically mounted cameras
observing the movement of zebrafish in shallow waters. The videos were then
processed using the automated tracking program TRex, which returns individual
files for each tracked fish for each video. Here, we are taking these individual
fish files and 1) converting the frame-by-frame data into summary statistics
for each video and 2) combining the data from the 500+ files into a singular
file. This file is used to export social data. See data_conversion_movement.Rmd
for the conversion of movement data. If using one file after the other make sure
to clear the working environment as many variables are shared across files.


## Setup

For our experiments fish occurred in either 3- or 4-fish groups. Change the
value of "fish" in the following chunk to account for the number of fish in the
experiment currently being analyzed.

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

#setting the working directory
knitr::opts_knit$set(root.dir = "D://2 Day Mini Kaolin Turb Flux Experiment/testing")
library(tidyverse)
library(data.table)
library(zoo)
library(sf)
#the number of fish per group
fish = 4
```

Setting up the code to be able to read in all of the files from a given
directory. We are also changing the "Inf" values TRex uses in place of NAs to
NAs and creating a master data frame to contain all of the summary statistics.

```{r file read in, message = FALSE}
#reading the file names into a list
list_csv <- list.files(pattern = "*.csv")
#from the list, creating a list of dataframes while excluding unused columns
masterlist <- lapply(list_csv, read_csv, col_select = c(3:4, 6:8))
#replacing the inf TRex uses in place of NAs with NAs
Inf2NA <- function(v) replace(v, is.infinite(v), NA)
fishlist <- lapply(masterlist, function(d) replace(d, TRUE, sapply(d, Inf2NA)))
finalframe <- data.frame(matrix(ncol = 7, nrow = (length(list_csv)/fish)))
colnames(finalframe) <- c("tankID", "recording", "nearest_neighbor",
                          "farthest_neighbor", "shoal_area", "shoal_area_var",
                          "polarity")
```

## Processing Files

The below code uses nested for-loops to process each of the datafiles and
eventually combine them into a single master file. For each video, TRex creates
a datafile for each detected fish- in the case of these experiments, that means
three or four fish per video (depending on the experiment). For analysis we are
lumping all fish behaviors across a group because individual identification is
impossible.

In this file we are extracting social data. We do this by... something about
X,Y coordinates and angles.

```{r loop, include = FALSE, message = FALSE}
i = 1
for (k in 1:(length(list_csv)/fish)) {  #across the whole file list
  #combining the files from all fish per video to be analyzed collectively
  shoal <- fishlist[i:(i+fish-1)]
  
  
  ##standardizing the length of each fish's file in a video so all files start at
  ##frame 0 and end at the last frame of the longest file
  #determining the last frame of a video by taking the largest frame value
  #from all of the fish datafiles
  maxframe <- max(unlist(lapply(shoal, function(m) m[,"frame"])))
  #creating a dataframe filled with NAs to pad the missing frames from each file
  dfNA <- data.frame(frame = 0:maxframe, fish_detected = NA, angle = NA,
                     X = NA, Y = NA)
  
  for (z in 1:fish) { #across one video of files
    
    #finding the frame the fish datafile starts on
    minfish <- min(shoal[[z]][, "frame"])
    
    #adding NA rows from dfNA to start of file when it starts on a frame later
    #than 0
    
    if (minfish > 0) {
      #across one fish file within a video
      shoal[[z]] <- rbind(dfNA[1:(minfish), ], shoal[[z]])
    }
    
    #finding the frame the fish datafile ends on
    maxfish <- max(shoal[[z]][, "frame"])
    
    #adding NA rows from the dfNA to the end of a file when it ends earlier than
    #the longest fish file from the video
    if (maxfish < maxframe) {
      #across one fish file within one video
      shoal[[z]] <- rbind(shoal[[z]], dfNA[(maxfish + 2):(maxframe + 1), ])
    }
  }  #z loop
    
  #creating an empty dataframe to hold the neighbor distance values
  distances <- data.frame(matrix(nrow = nrow(shoal[[1]]), ncol = 6))
  colnames(distances) <- c(1:6)
  #creating an empty dataframe to hold the polarity and shoal area values
  shoal.angle <- data.frame(matrix(nrow = nrow(shoal[[1]]), ncol = 2))
  colnames(shoal.angle) <- c("shoal_area", "polarity")
  
  
  #calculating nearest and farthest neighbor, polarity and shoal area
  for (a in 1:(maxframe + 1)) {
    #across one video of files
    
    #pulling the X,Y coordinates from each of the fish files in a video
    coor <- data.frame(matrix(nrow = fish, ncol = 2))
    colnames(coor) <- c("X", "Y")
    for (p in 1:fish) {
      coor$X[p] <- shoal[[p]][a, "X"]
      coor$Y[p] <- shoal[[p]][a, "Y"]
    }
    
    #calculating the distances between every fish
    neighbors <- as.vector(dist(coor))
    
    #adding the distance values into the distance dataframe
    for (d in 1:length(neighbors)) {
      distances[a, d] <- neighbors[d]
    }
    
    
    #calculating shoal areas, but only if all fish in the group are detected
    #for that frame
    if (anyNA(as.vector(unlist(coor))) == FALSE) {
      #sf needs for the first and last coordinates of a polygon to be the same
      #to "complete" the polygon
      coorpoly <- rbind(coor, coor[1,])
      #changing the classes of the columns back to numeric from list
      coorpoly$X <- unlist(coorpoly$X)
      coorpoly$Y <- unlist(coorpoly$Y)
      #converting the coordinates to a polygon and adding to dataframe
      shoal.angle$shoal_area[a] <- st_area(st_polygon(
                                   x = list(cbind(coorpoly$X, coorpoly$Y))))
    }
    
    
    #calculating polarity, which is the variance of fish angles
      #only calculates when all fish are detected
    shoal.angle$polarity[a] <- var(unlist(lapply(shoal,
                                                 function(v) v[a,"angle"])))
  }  #a loop
  
    #put the rolling average calculations here
  
  #creating an empty dataframe to hold the rolling averages of neighbor
  #distances
  rolldist <- data.frame(matrix(nrow = nrow(shoal[[1]]), ncol = 6))
  colnames(rolldist) <- c(1:6)
  #calculating the rolling averages for neighbor distances
  for (r in 1:6) {
    rolldist[r] <- rollapply(distances[,r], width = 7, FUN = function(x)
                             mean(x, na.rm = TRUE), partial = TRUE, fill = NA)
  }
    
  #creating an empty dataframe to hold the rolling averages for polarity and
  #shoal area
  roll.polarea <- data.frame(matrix(nrow = nrow(shoal[[1]]), ncol = 2))
  colnames(roll.polarea) <- c("shoal_area", "polarity")
  #calculating the rolling mean for shoal area
  roll.polarea$shoal_area <- rollapply(shoal.angle$shoal_area, width = 7,
                                       FUN = function(x) mean(x, na.rm = TRUE),
                                       partial = TRUE, fill = NA)
  #calculating the rolling mean for polarity
  roll.polarea$polarity <- rollapply(shoal.angle$polarity, width = 7,
                                       FUN = function(x) mean(x, na.rm = TRUE),
                                       partial = TRUE, fill = NA)
  
  #still need to pull out the max and min neighbor distances, but everything
  #else seems to be working as expected (I hope)
  
  str_hold <- strsplit(list_csv[i], split = "_")
  
  finalframe$tankID[k] <- str_hold[[1]][1]
  finalframe$recording[k] <- str_hold[[1]][2]
  finalframe$nearest_neighbor[k] <- min(unlist(rolldist, use.names = FALSE),
                                     na.rm = TRUE)
  finalframe$farthest_neighbor[k] <- max(unlist(rolldist, use.names = FALSE),
                                      na.rm = TRUE)
  finalframe$shoal_area[k] <- mean(roll.polarea$shoal_area, na.rm = TRUE)
  #this value seems wrong...
  finalframe$shoal_area_var[k] <- var(roll.polarea$shoal_area, na.rm = TRUE)
  finalframe$polarity[k] <- mean(roll.polarea$polarity, na.rm = TRUE)
  
  i = i+fish
}  #k loop
```