---
title: "Fluctuating Turbidity Data Conversion"
author: "Hannah Anderson"
date: "2023-01-16"
output: pdf_document
---
## Introduction

Data for this experiment was captured using vertically mounted cameras
observing the movement of zebrafish in shallow waters. The videos were then
processed using the automated tracking program TRex, which returns individual
files for each tracked fish for each video. Here, we are taking these individual
fish files and 1) converting the frame-by-frame data into summary statistics
for each video and 2) combining the data from the 500+ files into a singular
file. This file is used to export movement data. See [insert filename] for
the conversion of social and shoaling data. If using one file after the other
make sure to clear the working environment as many variables are shared across
files.


## Setup

For our experiments fish occurred in either 3- or 4-fish groups. Change the
value of "fish" in the following chunk to account for the number of fish in the
experiment currently being analyzed.

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

#setting the working directory
knitr::opts_knit$set(root.dir = "D://2 Day Mini Kaolin Turb Flux Experiment/testing")
library(tidyverse)
library(data.table)
library(zoo)
#the number of fish per group
fish = 4
```

Setting up the code to be able to read in all of the files from a given
directory. We are also changing the "Inf" values TRex uses in place of NAs to
NAs and creating a master data frame to contain all of the summary statistics.

The definition of each variable in the original datafile is as follows:
"tankID" = The ID of the tank chamber.In total there were eight 20-gallon tanks
           divided into three evenly sized chambers each for a total of 24
           tank chambers. Each of these tank chambers was fully isolated from
           neighboring chambers so that no water passed through and shared walls
           were opaque white. The IDs of the tank chambers occur in an X-Y
           format, where X is the number of the overall tank and Y is the
           individual chamber within the tank.
"recording" = The recording number for that particular video. Recordings
               started at Recording 1 and ended with Recording 9. Recording 6
               was excluded from analysis due to being an unplanned extra set
               of recordings.
"mean_speed" = The mean speed of the group's collective movements. Excludes
               frames where the fish was not detected.
"activity_ratio" = The ratio of frames the fish collectively spent active versus
                   inactive over the course of the video. Fish were defined as
                   "inactive" if their speed was below XXX and "active" if their
                   speed was XXX and above. Excludes frames where the fish was
                   not detected.
"speed_variance" = The variance of the speed of collectively all of the fish
                   over the course of a single recording. Excludes frames
                   where the fish was not detected.
"max_speed" = The maximum speed any fish in a group reached over the course of
              a recording.

```{r file read in, message = FALSE}
#reading the file names into a list
list_csv <- list.files(pattern = "*.csv")
#from the list, creating a list of dataframes while excluding unused columns
masterlist <- lapply(list_csv, read_csv, col_select = c(3:4, 6:8))
#replacing the inf TRex uses in place of NAs with NAs
Inf2NA <- function(v) replace(v, is.infinite(v), NA)
fishlist <- lapply(masterlist, function(d) replace(d, TRUE, sapply(d, Inf2NA)))
#creating a dataframe to hold the final product
finalframe <- data.frame(matrix(ncol = 7, nrow = (length(list_csv)/fish)))
colnames(finalframe) <- c("tankID", "recording", "mean_speed", "activity_ratio",
                          "speed_variance", "max_speed", "acceleration")
```

## Processing Files

The below code uses nested for-loops to process each of the datafiles and
eventually combine them into a single master file. For each video, TRex creates
a datafile for each detected fish- in the case of these experiments, that means
three or four fish per video (depending on the experiment). For analysis we are
lumping all fish behaviors across a group because individual identification is
impossible.

In this file we are extracting movement data. We do this by first processing the
X,Y coordinates of each fish into distances moved between frames, and from there
we calculate speed and then acceleration. To correct for noise in the tracking
process speed is calculated using a rolling average, with acceleration
calculated from the corrected speed values.

Once speed and acceleration have been calculated for each individual fish file
they are combined and summarized into a master datafile containing information
from all fish videos.

```{r loop, include = FALSE, message = FALSE}
i = 1
for (k in 1:(length(list_csv)/fish)) {  #across the whole file list
  #combining the files from all fish per video to be analyzed collectively
  shoal <- fishlist[i:(fish*i)]
  i = i+1
  
  
  ##standardizing the length of each fish's file in a video so all files start at
  ##frame 0 and end at the last frame of the longest file
  #determining the last frame of a video by taking the largest frame value
  #from all of the fish datafiles
  maxframe <- max(shoal[[1]][,"frame"])
  #creating a dataframe filled with NAs to pad the missing frames from each file
  dfNA <- data.frame(frame = 0:maxframe, fish_detected = NA, angle = NA,
                     X = NA, Y = NA)
  
  #creating a blank dataframe to hold the processed fish files from a video as
  #they are completed
  tempshoal <- data.frame(matrix(ncol = 2, nrow = 0))
  colnames(tempshoal) <- c("time", "speed")
  
  for (z in 1:fish) {  #across one video of files
    
    #finding the frame the fish datafile starts on
    minfish <- min(shoal[[z]][,"frame"])
    
    #adding NA rows from dfNA to start of file when it starts on a frame later
    #than 0
   
     if (minfish > 0) {  #across one fish file within a video
      shoal[[z]] <- rbind(dfNA[1:(minfish),], shoal[[z]])
    }
    
    #finding the frame the fish datafile ends on
    maxfish <- max(shoal[[z]][,"frame"])
    
    #adding NA rows from the dfNA to the end of a file when it ends earlier than
    #the longest fish file from the video
    if (maxfish < maxframe) {  #across one fish file within one video
      shoal[[z]] <- rbind(shoal[[z]], dfNA[(maxfish+2):(maxframe+1),])
    }
    

    #calculating the base distance moved between frames with no correcting
    dist <- data.frame(matrix(nrow = nrow(shoal[[z]]), ncol = 2))
    colnames(dist) <- c("frame", "distance_cm")
    dist$frame <- shoal[[z]][,"frame"]
    for (t in 1:(maxframe-1)) {  #within one fish's file
      dist$distance_cm[t+1] <- sqrt(
                    ((shoal[[z]][(t+1),"X"]) - (shoal[[z]][t,"X"]))^2 +
                    ((shoal[[z]][(t+1),"Y"]) - (shoal[[z]][t,"Y"]))^2)
    }
    #not sure why, but R has classified "distance_cm" as a list, so converting
    #it to a usable numeric class here
    dist$distance_cm <- as.numeric(dist$distance_cm)
    
    #calculating the base speed of the fish each second with no correcting
    simpspeed <- data.frame(matrix(nrow = floor(nrow(dist)/30), ncol = 2))
    colnames(simpspeed) <- c("time", "speed")
    simpspeed$time <- 1:(floor(nrow(dist)/30))
    #videos are ~30 frames per second (29.7 frames), using floor() to round down
    #to the nearest whole second
    n = 1
    for (s in 1:(floor(nrow(dist)/30))) {  #within one fish's file
      simpspeed$speed[s] <- sum(dist$distance_cm[n:(s*30)], na.rm = TRUE)
      n = n + 30
    }
  
    #correcting for noise in fish detection by calculating a rolling average speed
    rollspeed <- data.frame(matrix(nrow = floor(maxframe/30), ncol = 3))
    colnames(rollspeed) <- c("time", "speed", "acceleration")
    rollspeed$time <- 1:(floor(nrow(dist)/30))
    #using the rollapply() function from the package "zoo"
    #rollapply() handles NAs better than rollmean()
    #using the default centered alignment of the rolling window
    #excluding NAs from the averaging but including them as time points in the
      #window
    rollspeed$speed <- rollapply(simpspeed$speed, width = 7, FUN = function(x)
                                 mean(x, na.rm = TRUE), partial = TRUE, fill = NA)
      #this window value is arbitrary, might be worth piloting to figure out a
      #better one
    
    for (a in 1:((nrow(rollspeed))-1)) {
      rollspeed$acceleration[a+1] <- rollspeed$speed[a+1] - rollspeed$speed[a]
    }
    
    
    #combining the now completed rolling mean speed to a temporary holding frame
    tempshoal <- merge(x = tempshoal, y = rollspeed, all = TRUE)
  
  }
  #taking the file name of the currently iterated file and splitting the name
  #into the three sections divided by "_". The first and the third sections of
  #each filename are the tank ID and the recording number, respectively
  str_hold <- strsplit(list_csv[i], split = "_")
  
  #calculating the summary statistics and placing them in the correct place
  #on the master dataframe
  finalframe$tankID[k] <- str_hold[[1]][1]
  finalframe$recording[k] <- str_hold[[1]][2]
  finalframe$mean_speed[k] <- mean(tempshoal$speed, na.rm = TRUE)
  finalframe$activity_ratio[k] <- (sum(tempshoal$speed >= 10, na.rm = TRUE)/
                                    sum(tempshoal < 10, na.rm = TRUE)) #these are place-holder values
  finalframe$speed_variance[k] <- var(tempshoal$speed, na.rm = TRUE)
  finalframe$max_speed[k] <- max(tempshoal$speed, na.rm = TRUE)
  finalframe$acceleration[k] <- mean(tempshoal$acceleration, na.rm = TRUE)
}

write_csv(finalframe, "D://2 Day Mini Kaolin Turb Flux Experiment/test.csv")
```
