---
title: "Extracting Fish Speeds for Piloting"
author: "Hannah Anderson"
date: "2023-02-09"
output: pdf_document
---

## Introduction

See data_conversion_social.Rmd or data_conversion_movement.Rmd for full intro.

This code is only to export files containing speed values that can be
cross-referenced with the original videos to calibrate above which speed fish
should be classified as "active" and below which they should be classified as
"inactive," which will be used to create a ratio of time spend inactive versus
active for each recording.


## Setup

For our experiments fish occurred in either 3- or 4-fish groups. Change the
value of "fish" in the following chunk to account for the number of fish in the
video being processed. Please also change the directories to correspond to
the appropriate filepaths you are using.

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

#setting the working directory
knitr::opts_knit$set(root.dir = "E://2 Day Mini Kaolin Turb Flux Experiment/pilotfile/current")
library(tidyverse)
library(zoo)

#the folder you want the exported file to go
enddir <- ("E://2 Day Mini Kaolin Turb Flux Experiment/pilotfile")

#the number of fish in a group
fish = 4
```

## Setup Continued

The below chunk sets up the files and dataframes to be used for the rest of the
script.

```{r file import, message = FALSE}
#reading the file names into a list
list_csv <- list.files(pattern = "*.csv")
masterlist <- lapply(list_csv, read_csv, col_select = c(3:4, 6:8))
#replacing the inf TRex uses in place of NAs with NAs
Inf2NA <- function(v) replace(v, is.infinite(v), NA)
fishlist <- lapply(masterlist, function(d) replace(d, TRUE, sapply(d, Inf2NA)))

#extracting the tank ID and recording number from the original file name to be
#used in naming the exported file from this code
str_hold <- strsplit(list_csv[1], split = "_")
#creating column names for each fish in the group
whichfish <- paste0("fish", 1:fish)

#creating the dataframes that will hold the final version of the data
finalframe <- data.frame(matrix(ncol = fish+1, nrow = 0))
colnames(finalframe) <- c("second", whichfish)
rollfinal <- finalframe
```

## Standardizing File Lengths

For each video, TRex creates a datafile for each detected fish- in the case of
these experiments, that means three or four fish per video (depending on the
experiment). However, each fish file starts at the first frame the fish is
detected and ends at the last frame the fish is detected rather than at the
beginning and end of the video. The below chunk standardizes the lengths of
the fish files so they are all the same by having them all start on the first
frame of the video (frame 0) and ending on the latest frame any fish was
detected.

```{r standardizing, include = FALSE, message = FALSE}
for (k in 1:(fish)) {
  shoal <- fishlist[1:(fish)]
  #determining the last frame of a video by taking the largest frame value
  #from all of the fish datafiles
  maxframe <- max(unlist(lapply(shoal, function(m) m[,"frame"])))
  #creating a dataframe filled with NAs to pad the missing frames from each file
  dfNA <- data.frame(frame = 0:maxframe, fish_detected = NA, angle = NA,
                     X = NA, Y = NA)
  
  for (z in 1:fish) {
     #finding the frame the fish datafile starts on
    minfish <- min(shoal[[z]][,"frame"])
    
    #adding NA rows from dfNA to start of file when it starts on a frame later
    #than 0
   
     if (minfish > 0) {  #across one fish file within a video
      shoal[[z]] <- rbind(dfNA[1:(minfish),], shoal[[z]])
    }
    
    #finding the frame the fish datafile ends on
    maxfish <- max(shoal[[z]][,"frame"])
    
    #adding NA rows from the dfNA to the end of a file when it ends earlier than
    #the longest fish file from the video
    if (maxfish < maxframe) {  #across one fish file within one video
      #+2 to avoid duplicating last row and account for first frame being 0,
      #+1 to account for first frame being 0
      shoal[[z]] <- rbind(shoal[[z]], dfNA[(maxfish+2):(maxframe+1),])
    }
  }  #for loop Z
}  #for loop k
```

## Speed Calculations

The below chunk calculates the speed of all fish in a video from the detected
fish coordinates provided by TRex.

```{r speed, include = FALSE, message = FALSE}
#creating an empty dataframe to hold the distance values
dista <- data.frame(matrix(nrow = nrow(shoal[[z]]), ncol = (fish+1)))
colnames(dista) <- c("frame", whichfish)
dista$frame <- shoal[[z]][,"frame"]
#calculating the distances moved between frames with no corrections
for (t in 1:(maxframe-1)) {  #within one fish's file
  for (f in 1:(length(whichfish))) {
    dista[(t+1),whichfish[f]] <- sqrt(
                          ((shoal[[f]][(t+1),"X"]) - (shoal[[f]][t,"X"]))^2 +
                          ((shoal[[f]][(t+1),"Y"]) - (shoal[[f]][t,"Y"]))^2)
  }  #for loop f
}  #for loop t
#R has classified these distance columns as lists, so unlisting here
dista[,2:(fish+1)] <- unlist(dista[,2:(fish+1)])


#calculating the base speed of the fish each second with no correcting
n = 1
for (s in 1:(floor(nrow(dista))/30)) {
  for (w in 1:fish) {
      finalframe[s,whichfish[w]] <- sum(dista[n:(s*30),whichfish[w]],
                                        na.rm = TRUE)
  }
}

#correcting for noise in the calculated fish speeds using rolling means
for (r in 1:fish) {
  rollfinal[,whichfish[r]] <- rollapply(finalframe[,whichfish[r]], width = 7,
                                        FUN = function(x) mean (x, na.rm = TRUE),
                                        partial = TRUE, fill = NA)
}

finalframe$second <- 1:(nrow(finalframe))
rollfinal$second <- 1:(nrow(finalframe))

write_csv(finalframe, file.path(enddir, paste0((paste("base_speed", str_hold[[1]][1],
                                        str_hold[[1]][2], sep = "_")), ".csv")))
write_csv(rollfinal, file.path(enddir, paste0((paste("rolling_speed", str_hold[[1]][1],
                                        str_hold[[1]][2], sep = "_")), ".csv")))
```

