---
title: "Statistical Analyses of TRex Social Data"
author: "Hannah Anderson"
date: "2023-02-27"
output: pdf_document
editor_options: 
  chunk_output_type: console
---

## Introduction

This file was created to analyze data collected during a series of experiments
conducted from March 30 to August 14, 2022 on the effects of fluctuating
visibility on the behavior of zebrafish (Danio rerio). All data analyzed in this
file was collected using the automated animal tracking system TRex
(Walter & Couzin, 2021) and focuses on the social behaviors. For more
information on the experiments see the readme file located in this repository.

The specific measures analyzed here are shoal area, mean neighbor distance and
polarity. One value was calculated for each measure for each recording.
Shoal area was calculated by drawing a line between the midpoints of
each fish to create a polygon and then calculating the area in centimeters.
Mean neighbor distance was calculated by taking the distance between the
midpoints of each fish and then calculating the mean in centimeters. Polarity
was calculated by taking the circular variance of the angles of each fish and
then subtracting this value from one. Polarity is a value from 0-1 where values
closer to 1 indicate the fish are oriented in similar directions while values
closer to 0 indicate fish are oriented in highly different directions.

Mean shoal area and mean neighbor distance were analyzed using linear mixed
models with tank ID as a random factor to account for multiple recordings of
each tank. Because polarity is 0-1 bounded it has a non-normal distribution so
we analyzed it using a generalized linear mixed model with a beta family
distribution and tank ID as a random factor. All model residuals were visually
inspected for model assumptions. When necessary, data was log-10 transformed
to meet model assumptions.


## Setup

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

setwd("C://Users/hande/Documents/R/fluctuating-turbidity")
library(tidyverse)
library(glmmTMB)
library(emmeans)
library(car)
library(DHARMa)
library(data.table)

#importing the data
#chronic_clay <- read_csv("2week_clay_social_data.csv")
acute_clay <- read_csv("acute_clay_social_data.csv")
dye_stir <- read_csv("stir_dye_social_data.csv")
dye_nostir <- read_csv("nostir_dye_social_data.csv")

#changing the tank IDs to shoal IDs to ensure IDs are different between the
#experiments that get statistically compared
acute_clay <- acute_clay %>%
              rename_at("tankID", ~"shoalID")
acute_clay$shoalID <- gsub("tank", "clay", acute_clay$shoalID)
dye_stir <- dye_stir %>%
            rename_at("tankID", ~"shoalID")
dye_stir$shoalID <- gsub("tank", "dyeStir", dye_stir$shoalID)
dye_nostir <- dye_nostir %>%
              rename_at("tankID", ~"shoalID")
dye_nostir$shoalID <- gsub("tank", "dyeNoStir", dye_nostir$shoalID)

#adding a column to identify experiment
acute_clay$experiment <- "clay"
dye_stir$experiment <- "dye_stir"
dye_nostir$experiment <- "dye_no_stir"

#combining the clay and stirred dye data together in a single data frame for
#statistical comparison
clay_dye <- rbind(acute_clay, dye_stir)
#combining the stirred and unstirred dye data together in a single data frame
#for statistical comparison
stir_nostir <- rbind(dye_stir, dye_nostir)
```


## Analyses for Experiment 1: Chronic (Two Week) Clay Exposure
Coming back to this one later

```{r chronic clay}
#models including all recordings
chron.shoal <- glmmTMB(shoal_area ~ treatment*sex + (1|tankID),
                       data = chronic.clay, family = gaussian())
chron.neigh <- update(chron.shoal, mean_neighbor ~ .)
chron.pol <- update(chron.shoal, polarity ~ ., family = beta_family())

#checking the model assumptions
plot(simulateResiduals(fittedModel = chron.shoal)) #assumptions not met
plot(simulateResiduals(fittedModel = chron.neigh))
plot(simulateResiduals(fittedModel = chron.pol)) #visually seems acceptable

#transforming the data to meet model assumptions
chronic.trans <- chronic.clay[,c(1:6, 10)]
#chronic.log$polarity <- log(chronic.log$polarity)
chronic.trans$shoal_area <- log(chronic.clay$shoal_area)

#the transformed model
chrotrans.shoal <- update(chron.shoal, . ~ ., data = chronic.trans)

#checking the transformed model assumptions
plot(simulateResiduals(fittedModel = chrotrans.shoal))

#analyzing the results
Anova(chrotrans.shoal)
Anova(chron.neigh)
Anova(chron.pol)


#models including only recordings where the flux condition visibility was high
chron.shoal.high <- update(chron.shoal, . ~ ., data = high.chronic)
chron.neigh.high <- update(chron.neigh, . ~ ., data = high.chronic)
chron.pol.high <- update(chron.pol, . ~ ., data = high.chronic)

#checking the model assumptions
plot(simulateResiduals(fittedModel = chron.shoal.high)) #assumptions not met
plot(simulateResiduals(fittedModel = chron.neigh.high))
plot(simulateResiduals(fittedModel = chron.pol.high)) #visually seems acceptable

#transforming the data to meet model assumptions
chrohigh.trans <- high.chronic[,c(1:6, 10)]
chrohigh.trans$shoal_area <- log(high.chronic$shoal_area)

#the transformed model
chrolog.shoal.high <- update(chron.shoal, . ~ ., data = chrohigh.trans)

#checking the transformed model assumptions
plot(simulateResiduals(fittedModel = chrolog.shoal.high))

#analyzing the results
Anova(chrolog.shoal.high)
Anova(chron.neigh.high)
Anova(chron.pol.high)


#models including only recordings where the flux condition visibility was low
chron.shoal.low <- update(chron.shoal, . ~ ., data = low.chronic)
chron.neigh.low <- update(chron.neigh, . ~ ., data = low.chronic)
chron.pol.low <- update(chron.pol, . ~ ., data = low.chronic)

#checking the model assumptions
plot(simulateResiduals(fittedModel = chron.shoal.low)) #assumptions not met
plot(simulateResiduals(fittedModel = chron.neigh.low))
plot(simulateResiduals(fittedModel = chron.pol.low)) #visually seems acceptable

#transforming the data to meet model assumptions
chrolow.trans <- low.chronic[,1:6, 10]
chrolow.trans$shoal_area <- log(low.chronic$shoal_area)

#the transformed model
chrolog.shoal.low <- update(chron.shoal, . ~ ., data = chrolow.trans)

#checking the transformed model assumptions
plot(simulateResiduals(fittedModel = chrolog.shoal.low))

#analyzing the models
Anova(chrolog.shoal.low)
Anova(chron.neigh.low)
Anova(chron.pol.low)
```


## Direct Comparisons

For the direct comparison models we compared the response variables of the compared experiments during the recording on the second day where the flux and stable conditions had the same visibility (recording 4).

The first set of direct comparisons is between the acute (two day) clay experiment and the stirred dye experiment to explore any potential differences between clay- and dye-imparied visibility. The second set of direct comparisons is between the stirred and unstirred dye experiments to determine if there was an effect of stirring on fish response to the treatments.

```{r direct}
#sub-setting the data to only include recording 4
direct_claydye <- clay_dye %>%
                  filter(recording == "rec4")
direct_dye <- stir_nostir %>%
              filter(recording == "rec4")


#the clay vs dye models

#mean shoal area
d_claydye_shoal <- lm(shoal_area ~ treatment + experiment,
                      data = direct_claydye)
#mean neighbor distance
d_claydye_neigh <- update(d_claydye_shoal, mean_neighbor ~ .)
#polarity
d_claydye_pol <- update(d_claydye_shoal, polarity ~ .)

#checking the model assumptions
plot(d_claydye_shoal)
plot(d_claydye_neigh)
plot(d_claydye_pol) #assumptions not met

#transforming the data to meet model assumptions
d_trans_claydye <- direct_claydye
d_trans_claydye$polarity <- asin(sqrt(d_trans_claydye$polarity))

#the transformed model
trans_d_claydye_pol <- update(d_claydye_pol, . ~ ., data = d_trans_claydye)

#checking the transformed model assumptions
plot(trans_d_claydye_pol)

#analyzing the models
Anova(d_claydye_shoal)
Anova(d_claydye_neigh)
Anova(trans_d_claydye_pol)


#the stirred vs unstirred dye models

#mean shoal area
d_dyedye_shoal <- lm(shoal_area ~ treatment*experiment, data = direct_dye)
#mean neighbor distance
d_dyedye_neigh <- update(d_dyedye_shoal, mean_neighbor ~ .)
#polarity
d_dyedye_pol <- update(d_dyedye_shoal, polarity ~ .)

#checking the model assumptions
plot(d_dyedye_shoal)
plot(d_dyedye_neigh)
plot(d_dyedye_pol)

#analyzing the models
Anova(d_dyedye_shoal)
Anova(d_dyedye_neigh)
Anova(d_trans_dye_pol)
```


## Full Model Comparisons of the Acute Experiments

```{r full}
#averaging together the two morning recordings for each measure and for each
#experiment by comparison (clay/dye, dye stirred/dye unstirred)
morn_claydye <- clay_dye %>%
                filter(recording %in% c("rec1", "rec3")) %>%
                select(-recording) %>%
                group_by(shoalID, treatment, experiment) %>%
                summarize(shoal_area = mean(shoal_area),
                          mean_neighbor = mean(mean_neighbor),
                          polarity = mean(polarity))
morn_dyedye <- stir_nostir %>%
               filter(recording %in% c("rec1", "rec3")) %>%
               select(-recording) %>%
               group_by(shoalID, treatment, experiment) %>%
               summarize(shoal_area = mean(shoal_area),
                         mean_neighbor = mean(mean_neighbor),
                         polarity = mean(polarity))
#labeling these values as the morning values
morn_claydye$time <- "morning"
morn_dyedye$time <- "morning"
#averaging together the afternoon values
afte_claydye <- clay_dye %>%
                filter(recording %in% c("rec2", "rec4")) %>%
                select(-recording) %>%
                group_by(shoalID, treatment, experiment) %>%
                summarize(shoal_area = mean(shoal_area),
                          mean_neighbor = mean(mean_neighbor),
                          polarity = mean(polarity))
afte_dyedye <- stir_nostir %>%
               filter(recording %in% c("rec2", "rec4")) %>%
               select(-recording) %>%
               group_by(shoalID, treatment, experiment) %>%
               summarize(shoal_area = mean(shoal_area),
                         mean_neighbor = mean(mean_neighbor),
                         polarity = mean(polarity))
#labeling these values as the afternoon values
afte_claydye$time <- "afternoon"
afte_dyedye$time <- "afternoon"
#combining the averaged morning and afternoon values into a single dataframe
#for each two-experiment comparison
full_claydye <- rbind(morn_claydye, afte_claydye)
full_dyedye <- rbind(morn_dyedye, afte_dyedye)


#the clay vs dye models

#mean shoal area
claydye_shoal <- lm(shoal_area ~ treatment + experiment + time,
                    data = full_claydye)
#mean neighbor distance
claydye_neigh <- update(claydye_shoal, mean_neighbor ~ .)
#polarity
claydye_pol <- update(claydye_shoal, polarity ~ .)

#checking the model assumptions
plot(claydye_shoal) #assumptions not met
plot(claydye_neigh)
plot(claydye_pol)

#transforming the data to meet model assumptions
trans_full_claydye <- full_claydye
trans_full_claydye$shoal_area <- log(trans_full_claydye$shoal_area)

#the transformed model
trans_claydye_shoal <- update(claydye_shoal, . ~ ., data = trans_full_claydye)

#checking the transformed model assumptions
plot(trans_claydye_shoal)

#analyzing the models
Anova(trans_claydye_shoal)
Anova(claydye_neigh)
Anova(claydye_pol)


#the stirred vs unstirred dye models

#mean shoal area
dyedye_shoal <- lm(shoal_area ~ treatment*experiment*time, data = full_dyedye)
#mean neighbor distance
dyedye_neigh <- update(dyedye_shoal, mean_neighbor ~ .)
#polarity
dyedye_pol <- update(dyedye_shoal, polarity ~ .)

#checking the model assumptions
plot(dyedye_shoal)
plot(dyedye_neigh)
plot(dyedye_pol) #assumptions not met

#transforming the data to meet model assumptions
trans_full_dyedye <- full_dyedye
trans_full_dyedye$polarity <- asin(sqrt(trans_full_dyedye$polarity))

#the transformed model
trans_dyedye_pol <- update(dyedye_pol, . ~ ., data = trans_full_dyedye)

#checking the transformed model assumptions
plot(trans_dyedye_pol)

#analyzing the models
Anova(dyedye_shoal)
Anova(dyedye_neigh)
Anova(trans_dyedye_pol)
```