---
title: "Statistical Analyses of TRex Social Data"
author: "Hannah Anderson"
date: "2023-02-27"
output: pdf_document
editor_options: 
  chunk_output_type: console
---

## Introduction

This file was created to analyze data collected during a series of experiments
conducted from March 30 to August 14, 2022 on the effects of fluctuating
visibility on the behavior of zebrafish (Danio rerio). All data analyzed in this
file was collected using the automated animal tracking system TRex
(Walter & Couzin, 2021) and focuses on the social behaviors. For more
information on the experiments see the readme file located in this repository.

The specific measures analyzed here are shoal area, mean neighbor distance and
polarity. One value was calculated for each measure for each recording.
Shoal area was calculated by drawing a line between the midpoints of
each fish to create a polygon and then calculating the area in centimeters.
Mean neighbor distance was calculated by taking the distance between the
midpoints of each fish and then calculating the mean in centimeters. Polarity
was calculated by taking the circular variance of the angles of each fish and
then subtracting this value from one. Polarity is a value from 0-1 where values
closer to 1 indicate the fish are oriented in similar directions while values
closer to 0 indicate fish are oriented in highly different directions.

Mean shoal area and mean neighbor distance were analyzed using linear mixed
models with tank ID as a random factor to account for multiple recordings of
each tank. Because polarity is 0-1 bounded it has a non-normal distribution so
we analyzed it using a generalized linear mixed model with a beta family
distribution and tank ID as a random factor. All model residuals were visually
inspected for model assumptions. When necessary, data was log-10 transformed
to meet model assumptions.


## Setup

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

setwd("C://Users/hande/Documents/R/fluctuating-turbidity")
library(tidyverse)
library(glmmTMB)
library(emmeans)
library(car)
library(lmtest)
library(DHARMa)

#importing the data
chronic.clay <- read_csv("2week_clay_social_data.csv")
acute.clay <- read_csv("2day_clay_social_data.csv")
dye.stir <- read_csv("dye_stir_social_data.csv")
dye.nostir <- read_csv("dye_nostir_social_data.csv")

#subsetting the data by the visibility conditions in the flux condition
##high visibility
high.chronic <- chronic.clay %>%
                filter(recording %in% c("rec3", "rec4", "rec7", "rec9"))
high.acute <- acute.clay %>%
              filter(recording %in% c("rec2", "rec4"))
high.stir <- dye.stir %>%
             filter(recording %in% c("rec2", "rec4"))
high.nostir <- dye.nostir %>%
               filter(recording %in% c("rec2", "rec4"))
#low visibility
low.chronic <- chronic.clay %>%
               filter(recording %in% c("rec1", "rec2", "rec5", "rec8"))
low.acute <- acute.clay %>%
             filter(recording %in% c("rec1", "rec3"))
low.stir <- dye.stir %>%
            filter(recording %in% c("rec1", "rec3"))
low.nostir <- dye.nostir %>%
              filter(recording %in% c("rec1", "rec3"))
```


## Analyses for Experiment 1: Chronic (Two Week) Clay Exposure

```{r chronic clay}
#models including all recordings
chron.shoal <- glmmTMB(shoal_area ~ treatment*sex + (1|tankID),
                       data = chronic.clay, family = gaussian())
chron.neigh <- update(chron.shoal, mean_neighbor ~ .)
chron.pol <- update(chron.shoal, polarity ~ ., family = beta_family())

#checking the model assumptions
plot(simulateResiduals(fittedModel = chron.shoal)) #assumptions not met
plot(simulateResiduals(fittedModel = chron.neigh))
plot(simulateResiduals(fittedModel = chron.pol)) #visually seems acceptable

#transforming the data to meet model assumptions
chronic.trans <- chronic.clay[,c(1:6, 10)]
#chronic.log$polarity <- log(chronic.log$polarity)
chronic.trans$shoal_area <- log(chronic.clay$shoal_area)

#the transformed model
chrotrans.shoal <- update(chron.shoal, . ~ ., data = chronic.trans)

#checking the transformed model assumptions
plot(simulateResiduals(fittedModel = chrotrans.shoal))

#analyzing the results
Anova(chrotrans.shoal)
Anova(chron.neigh)
Anova(chron.pol)


#models including only recordings where the flux condition visibility was high
chron.shoal.high <- update(chron.shoal, . ~ ., data = high.chronic)
chron.neigh.high <- update(chron.neigh, . ~ ., data = high.chronic)
chron.pol.high <- update(chron.pol, . ~ ., data = high.chronic)

#checking the model assumptions
plot(simulateResiduals(fittedModel = chron.shoal.high)) #assumptions not met
plot(simulateResiduals(fittedModel = chron.neigh.high))
plot(simulateResiduals(fittedModel = chron.pol.high)) #visually seems acceptable

#transforming the data to meet model assumptions
chrohigh.trans <- high.chronic[,c(1:6, 10)]
chrohigh.trans$shoal_area <- log(high.chronic$shoal_area)

#the transformed model
chrolog.shoal.high <- update(chron.shoal, . ~ ., data = chrohigh.trans)

#checking the transformed model assumptions
plot(simulateResiduals(fittedModel = chrolog.shoal.high))

#analyzing the results
Anova(chrolog.shoal.high)
Anova(chron.neigh.high)
Anova(chron.pol.high)


#models including only recordings where the flux condition visibility was low
chron.shoal.low <- update(chron.shoal, . ~ ., data = low.chronic)
chron.neigh.low <- update(chron.neigh, . ~ ., data = low.chronic)
chron.pol.low <- update(chron.pol, . ~ ., data = low.chronic)

#checking the model assumptions
plot(simulateResiduals(fittedModel = chron.shoal.low)) #assumptions not met
plot(simulateResiduals(fittedModel = chron.neigh.low))
plot(simulateResiduals(fittedModel = chron.pol.low)) #visually seems acceptable

#transforming the data to meet model assumptions
chrolow.trans <- low.chronic[,1:6, 10]
chrolow.trans$shoal_area <- log(low.chronic$shoal_area)

#the transformed model
chrolog.shoal.low <- update(chron.shoal, . ~ ., data = chrolow.trans)

#checking the transformed model assumptions
plot(simulateResiduals(fittedModel = chrolog.shoal.low))

#analyzing the models
Anova(chrolog.shoal.low)
Anova(chron.neigh.low)
Anova(chron.pol.low)
```


## Analyses for Experiment 2: Acute (Two Day) Clay Exposure

```{r acute clay}
#models including all recordings
acute.shoal <- glmmTMB(shoal_area ~ treatment + (1|tankID),
                       data = acute.clay, family = gaussian())
acute.neigh <- update(acute.shoal, mean_neighbor ~ .)
acute.pol <- update(acute.shoal, polarity ~ ., family = beta_family())

#checking the model assumptions
plot(simulateResiduals(fittedModel = acute.shoal))
plot(simulateResiduals(fittedModel = acute.neigh))
plot(simulateResiduals(fittedModel = acute.pol))

#analyzing the models
Anova(acute.shoal)
Anova(acute.neigh)
Anova(acute.pol)


#models including only recordings where the flux condition visibility was high
acute.shoal.high <- update(acute.shoal, . ~ ., data = high.acute)
acute.neigh.high <- update(acute.neigh, . ~ ., data = high.acute)
acute.pol.high <- update(acute.pol, . ~ ., data = high.acute)

#checking the model assumptions
plot(simulateResiduals(fittedModel = acute.shoal.high))
plot(simulateResiduals(fittedModel = acute.neigh.high))
plot(simulateResiduals(fittedModel = acute.pol.high))

#analyzing the models
Anova(acute.shoal.high)
Anova(acute.neigh.high)
Anova(acute.pol.high)


#models including only recordings where the flux condition visibility was low
acute.shoal.low <- update(acute.shoal, . ~ ., data = low.acute)
acute.neigh.low <- update(acute.neigh, . ~ ., data = low.acute)
acute.pol.low <- update(acute.pol, . ~ ., data = low.acute)

#checking the model assumptions
plot(simulateResiduals(fittedModel = acute.shoal.low))
plot(simulateResiduals(fittedModel = acute.neigh.low))
plot(simulateResiduals(fittedModel = acute.pol.low))

#analyzing the models
Anova(acute.shoal.low)
Anova(acute.neigh.low)
Anova(acute.pol.low)

#post hoc pairwise comparison of the significant result
acute.pol.low %>% emmeans(pairwise ~ treatment)
```


## Analyses for Experiment 3: Acute (Two Day) Dye Exposure with Stirring

```{r dye stir}
#models including all recordings
stir.shoal <- glmmTMB(shoal_area ~ treatment + (1|tankID),
                       data = dye.stir, family = gaussian())
stir.neigh <- update(stir.shoal, mean_neighbor ~ .)
stir.pol <- update(stir.shoal, polarity ~ ., family = beta_family())

#checking the model assumptions
plot(simulateResiduals(fittedModel = stir.shoal)) #assumptions not met
plot(simulateResiduals(fittedModel = stir.neigh))
plot(simulateResiduals(fittedModel = stir.pol))

#transforming the data to meet model assumptions
stir.trans <- dye.stir[,c(1:5, 9)]
stir.trans$shoal_area <- log(dye.stir$shoal_area)

#the transformed model
stirlog.shoal <- update(stir.shoal, . ~ ., data = stir.trans)

#checking the transformed model assumptions
plot(simulateResiduals(fittedModel = stirlog.shoal))

#analyzing the models
Anova(stirlog.shoal)
Anova(stir.neigh)
Anova(stir.pol)


#models including only recordings where the flux condition visibility was high
stir.shoal.high <- update(stir.shoal, . ~ ., data = high.stir)
stir.neigh.high <- update(stir.neigh, . ~ ., data = high.stir)
stir.pol.high <- update(stir.pol, . ~ ., data = high.stir)

#checking the model assumptions
plot(simulateResiduals(fittedModel = stir.shoal.high))
plot(simulateResiduals(fittedModel = stir.neigh.high))
plot(simulateResiduals(fittedModel = stir.pol.high))

#analyzing the models
Anova(stir.shoal.high)
Anova(stir.neigh.high)
Anova(stir.pol.high)


#models including only recordings where the flux condition visibility was low
stir.shoal.low <- update(stir.shoal, . ~ ., data = low.stir)
stir.neigh.low <- update(stir.neigh, . ~ ., data = low.stir)
stir.pol.low <- update(stir.pol, . ~ ., data = low.stir)

#checking the model assumptions
plot(simulateResiduals(fittedModel = stir.shoal.low))
plot(simulateResiduals(fittedModel = stir.neigh.low))
plot(simulateResiduals(fittedModel = stir.pol.low))

#analyzing the models
Anova(stir.shoal.low)
Anova(stir.neigh.low)
Anova(stir.pol.low)
```



## Analyses for Experiment 4: Acute (Two Day) Dye Exposure without Stirring

```{r dye nostir}
#models including all recordings
nostir.shoal <- glmmTMB(shoal_area ~ treatment + (1|tankID),
                       data = dye.nostir, family = gaussian())
nostir.neigh <- update(nostir.shoal, mean_neighbor ~ .)
nostir.pol <- update(nostir.shoal, polarity ~ ., family = beta_family())

#checking the model assumptions
plot(simulateResiduals(fittedModel = nostir.shoal))
plot(simulateResiduals(fittedModel = nostir.neigh))
plot(simulateResiduals(fittedModel = nostir.pol))

#analyzing the models
Anova(nostir.shoal)
Anova(nostir.neigh)
Anova(nostir.pol)


#models including only recordings where the flux condition visibility was high
nostir.shoal.high <- update(nostir.shoal, . ~ ., data = high.nostir)
nostir.neigh.high <- update(nostir.neigh, . ~ ., data = high.nostir)
nostir.pol.high <- update(nostir.pol, . ~ ., data = high.nostir)

#checking the model assumptions
plot(simulateResiduals(fittedModel = nostir.shoal.high))
plot(simulateResiduals(fittedModel = nostir.neigh.high))
plot(simulateResiduals(fittedModel = nostir.pol.high))

#analyzing the models
Anova(nostir.shoal.high)
Anova(nostir.neigh.high)
Anova(nostir.pol.high)


#models including only recordings where the flux condition visibility was low
nostir.shoal.low <- update(nostir.shoal, . ~ ., data = low.nostir)
nostir.neigh.low <- update(nostir.neigh, . ~ ., data = low.nostir)
nostir.pol.low <- update(nostir.pol, . ~ ., data = low.nostir)

#checking the model assumptions
plot(simulateResiduals(fittedModel = nostir.shoal.low))
plot(simulateResiduals(fittedModel = nostir.neigh.low))
plot(simulateResiduals(fittedModel = nostir.pol.low))

#analyzing the models
Anova(nostir.shoal.low)
Anova(nostir.neigh.low)
Anova(nostir.pol.low)
```